---
title: "PSYCH308C - Data Analysis 03 (DA03)"
author: "Brady C. Jackson"
date: "2025/02/04"

# Write document output to HTML, Word, and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Prompt

You now work for a nationwide transitional living program. They serve homeless individuals between the age of 16-25.  
The various services they are providing include: (1) a service that helps youth get a job (2) a literacy program (3)  
a program to help youth graduate from high school and (4) shelter to live in. They've collected data on the youth in  
their program on income, literacy, and high school graduation (see below for an explanation of the variables). The   
program ultimately cares about what best predicts the successful transition of these youth. According to the   
literature, high school graduation and income explain successful transitioning best. The CEO thinks that from   
her experience literacy is actually an important factor to consider in addition to graduation and income.   
Test both the model based on literature and the CEO's proposed model to determine what best predicts successful  
transition. 

## Variables 

| Variable | Type | Description |
|:-----|:---|:------------------|
| `Income` | $ | Annual income |
| `Illiteracy` | 1-3 scale | Level of illiteracy on a scale of 0-3 (higher being more illiterate) |
| `HS.Grad` | Factor: <br>Reference = Did not Graduate, <br>D1 = On Time, <br>D2 = Later | Whether they graduated on time (18 years old), later (any age after 18), <br>or did not graduate at all. |
| `Success` | 1-10 scale | Successful transition scaled based on a variety of factors compiling to <br>an ultimate score between the values of 0-10 |


## Assignment

### Part 1

Conduct the appropriate analyses (including assumptions) and write a results section for an APA journal.   
Include tables and figures as necessary.

### Part 2
| 1. Briefly explain your findings to the CEO. What recommendation would you make for the program? 
| 2. Report the two different pieces of evidence that you can use to determine if high school graduation is a   
|    significant predictor of success. 
| 3. According to your best model, what is the slope of the line for graduated on time? What is the slope for   
|    did not graduate? Are they the same or are they different? Explain why or why not. 
| 4. The economic structure of Major League Baseball allows some teams to make substantially more money than others,   
|    which in turn allows some teams to spend much more on player salaries. These teams might therefore be expected to have   
|    better players and win more games as a result. Suppose that after collecting data on team payroll (in millions of   
|    dollars) and season win total for 2010, we find a regression equation of (Wins = 71.87 + 0.10Payroll - 0.06League),  
|    where League is a predictor variable that equals 0 if the team plays in the National League or 1 if the team plays   
|    in the American League. What is the intercept for a team in the American League? What is the slope?
    
---ONLY WRITE CODE BELOW THIS LINE---

## Code

### Libraries

Load all requisite libraries here.

```{r package_loading, message=FALSE, warning=FALSE}
# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(jmv)       # for descriptive
library(ggplot2)
library(dplyr)
library(corrplot)    # For fancy covariance matrix plots
library(apaTables)   # For Word formatted tables
library(car)         # for ncvTest (Breusch Pagan)
# library(stringr)   # for sub_str operations
library(psych)
# library(Hmisc)     # for fun.dat substitution
# library(see)       # for outliers analysis 
library(magrittr)
# library(AER)
library(rlang)
library(mvnTest)
library(easystats)
library(patchwork)
library(mvtnorm)
```


### Metadata

This section of code is to setup some general variables that we'll use throughout the code (e.g. figure colors, etc)

```{r metadata}
# First we'll defines some meta-data to use in all of our plots so they're nice and clean
font_color = "#4F81BD"
grid_color_major = "#9BB7D9"
grid_color_minor = "#C8D7EA"
back_color = "gray95"
rb_colmap = colorRampPalette( c("firebrick", "grey86", "dodgerblue3") )(200)

# I'm going to try to save off my preferred ggplot theme combinations as a unqiue theme object that I can just reference
# later in the code....totally unclear if ggplot works this way....
my_gg_theme = theme_minimal() +
              theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
                     axis.title.x = element_text(color = font_color),
                     axis.title.y = element_text(color = font_color),
                     axis.text.x = element_text(color = font_color),
                     axis.text.y = element_text(color = font_color),
                     legend.title = element_text(color = font_color),
                     legend.text = element_text(color = font_color),
                     panel.grid.minor = element_line(color = grid_color_minor),
                     panel.grid.major = element_line(color = grid_color_major),
                     panel.background = element_rect(fill = back_color, color = font_color)
                   )

```

### Load and View data

```{r load_raw_data}

# Load data as downloaded from Canvas
transition_dat = read.csv("./308C.Data.DA3.csv")

# Rename columns to lower because why not
colnames(transition_dat) <- tolower( colnames(transition_dat) )

# Create Dummy Code variables for graduation level. These will be named:

# We're going to use "Graduated ontime" as a reference so we can see the delta-effects on transition of
# 1) Whether graduating late results in any significant decrease in transition compared to graduating on time
# 2) Whether failing to graduate results in any significant decrease in transition compared to graduating on time. 

# NOTE: For comparison's sake we'll create Dummy codes that use Did Not Graduate for another model so we can also get
#       the "Graduate Late <-> Did not graduate" comparison
#

# Create columns for Dummy Codes referenced to: Graduated ontime
transition_dat$grad_late_refd_to_ontime <- transition_dat$hs.grad
transition_dat$grad_none_refd_to_ontime <- transition_dat$hs.grad

# Create columns for Dummy Codes referenced to: Did not graduate
transition_dat$grad_late_refd_to_none   <- transition_dat$hs.grad
transition_dat$grad_ontime_refd_to_none <- transition_dat$hs.grad

# Dummy code replacement logic for the OnTime Reference Case. Codes for this set are as follows:
# [ cond     | ontime | late | none ]
# [   ontime |      0 |    0 |    0 ]
# [   late   |      0 |    1 |    0 ]
# [   none   |      0 |    0 |    1 ]
transition_dat$grad_late_refd_to_ontime[transition_dat$hs.grad == 'Graduated ontime'] <- 0
transition_dat$grad_late_refd_to_ontime[transition_dat$hs.grad == 'Graduated later']  <- 1
transition_dat$grad_late_refd_to_ontime[transition_dat$hs.grad == 'Did not graduate'] <- 0

transition_dat$grad_none_refd_to_ontime[transition_dat$hs.grad == 'Graduated ontime'] <- 0
transition_dat$grad_none_refd_to_ontime[transition_dat$hs.grad == 'Graduated later']  <- 0
transition_dat$grad_none_refd_to_ontime[transition_dat$hs.grad == 'Did not graduate'] <- 1


# Dummy code replacement logic for the None Reference Case. Codes for this set are as follows:
# [ cond     | ontime | late | none ]
# [   none   |      0 |    0 |    0 ]
# [   late   |      0 |    1 |    0 ]
# [   ontime |      1 |    0 |    0 ]
transition_dat$grad_late_refd_to_none[transition_dat$hs.grad == 'Graduated ontime'] <- 0
transition_dat$grad_late_refd_to_none[transition_dat$hs.grad == 'Graduated later']  <- 1
transition_dat$grad_late_refd_to_none[transition_dat$hs.grad == 'Did not graduate'] <- 0

transition_dat$grad_ontime_refd_to_none[transition_dat$hs.grad == 'Graduated ontime'] <- 1
transition_dat$grad_ontime_refd_to_none[transition_dat$hs.grad == 'Graduated later']  <- 0
transition_dat$grad_ontime_refd_to_none[transition_dat$hs.grad == 'Did not graduate'] <- 0

# Now we need to convert all of our integer arrays to int types since we primed them as character arrays
transition_dat$grad_late_refd_to_ontime <- as.integer(transition_dat$grad_late_refd_to_ontime)
transition_dat$grad_none_refd_to_ontime <- as.integer(transition_dat$grad_none_refd_to_ontime)
transition_dat$grad_late_refd_to_none   <- as.integer(transition_dat$grad_late_refd_to_none)
transition_dat$grad_ontime_refd_to_none <- as.integer(transition_dat$grad_ontime_refd_to_none)

```

### Descriptive Statistics - Raw Data

This section will look at base descriptive statistics of the raw data to help identify data anomalies and check   
normality of predictor variables (break length - minutes)

```{r descriptive_stats_raw}

# Get JMV descriptives of raw success, professionalism, competence, and motivation,
# Tabulate mean, variance, skew, and kurtosis so we can check normality.
# Output histograms and qq plots so we can spot outliers and review distribution shape
#  We're going to split by graduation level so we can best understand data trends within the categorical variables.
# This will also help us see collinearity trends between predictors potentially.
transition_descr_splt = jmv::descriptives( transition_dat,
                                           vars = c('income', 'illiteracy', 'success'),
                                           splitBy = 'hs.grad',
                                           hist = TRUE,
                                           dens = TRUE,
                                           qq = TRUE,
                                           sd = TRUE,
                                           variance = TRUE,
                                           se = TRUE,
                                           skew = TRUE,
                                           kurt = TRUE,
                                           missing = TRUE)
print(transition_descr_splt)


# Now we'll look at descriptives of the unsplit set
cat( 
     paste(
           '------------------',
           '---- Combined ----',
           '------------------',
           sep = '\n'
     ) 
   )

transition_descr_comb = jmv::descriptives( transition_dat,
                                           vars = c('income', 'illiteracy', 'success', 'hs.grad'),
                                           hist = TRUE,
                                           dens = TRUE,
                                           qq = TRUE,
                                           sd = TRUE,
                                           variance = TRUE,
                                           se = TRUE,
                                           skew = TRUE,
                                           kurt = TRUE,
                                           missing = TRUE)
print(transition_descr_comb)

```

### Correlation Plots - Raw Data

Visualize the covariance matrix to understand correlation between Success and Professionalism and Competence (Model 1)
A second correlation matrix is also computed which includes Motivation (for model 2)


```{r correlation_plots_raw}

# Model 1 Correlation Tables
cat( 
     paste(
           '-----------------',
           '---- Model 1 ----',
           '-----------------',
           sep = '\n'
     ) 
   )

# Our first model will look at correlations associated with the simpler model:
# success ~ income + graduation
# But, since graduation is categorical, we strip it out.
m1_subset = transition_dat[ c('income', 'success') ]
transition_cor_m1 <- stats::cor( m1_subset )
corrplot( transition_cor_m1,
          method="color",
          type = "full",
          addCoef.col = "black",
          col = rb_colmap,
          tl.col = font_color )

# Print correlation tables (annoying that we have to use two different correlation functions, 1 for plotting and
# 1 for tables, but oh well)
transition_cor_tab_m1 <- jmv::corrMatrix(m1_subset, flag = TRUE)
print( transition_cor_tab_m1 )


# Model 2 Correlation Tables
cat( 
     paste(
           '-----------------',
           '---- Model 2 ----',
           '-----------------',
           sep = '\n'
     ) 
   )

# Our second model will look at correlations associated with the additional predictor (literacy) model:
# success ~ income + graduation + literacy
# But, since graduation is categorical, we strip it out.
m2_subset = transition_dat[ c('income', 'success', 'illiteracy') ]
transition_cor_m2 <- stats::cor( m2_subset )
corrplot( transition_cor_m2,
          method="color",
          type = "full",
          addCoef.col = "black",
          col = rb_colmap,
          tl.col = font_color )

# Print correlation tables (annoying that we have to use two different correlation functions, 1 for plotting and
# 1 for tables, but oh well)
transition_cor_tab_m2 <- jmv::corrMatrix(m2_subset, flag = TRUE)
print( transition_cor_tab_m2 )

```
### Scatterplots of All Predictor Vars

Since scatterplots are not variables based, we can create a scatterplot with Transition as a function of each of the  
2x continuous predictor vars. Scatterplots are broken out here, separate from residual plots and homoscedasticity  
testing for readability.

**NOTE:** I'm not using centered data for the scatterplots as it's unintuitive but I may add them later if necessary

```{r scatterplots_raw}

# Scatterplot 1: Transition (uncentered) vs. Income
success_scatter_tvinc <- ggplot(transition_dat, aes(income, success) )

success_scatter_tvinc +
    geom_point() +
    geom_smooth(method = "lm", colour = "Red") +
    ggtitle("Transition Success Relation to Student Income") +
    labs(y = "Transition Success (1-10)", x = "Student Income ($USD)") +
    my_gg_theme


# Scatterplot 2: Transition (uncentered) vs. Illiteracy
success_scatter_tvil <- ggplot(transition_dat, aes(illiteracy, success) )

success_scatter_tvil +
    geom_point() +
    geom_smooth(method = "lm", colour = "Red") +
    ggtitle("Transition Success Relation to Student Illiteracy") +
    labs(y = "Transition Success (1-10)", x = "Student Illiteracy (0-3)") +
    my_gg_theme
```

### Simple Linear Model Defintion

We need to create two simple linear models for various data visualization and analysis efforts. The following code   
creates these two models:

Model 1: transition ~ graduation + income  
Model 2: transition ~ graduation + income + illiteracy  

```{r simple_model_definition}
# We create the linear models using separate datasets simply for clarity of code and readability

# Graduation referenced to OnTime, w/ income only
transition_simp_mod_m1_ontime <- lm( success ~ income + grad_late_refd_to_ontime + grad_none_refd_to_ontime, 
                                     data = transition_dat)
# Graduation referenced to None, w/ income only
transition_simp_mod_m1_none  <- lm( success ~ income + grad_late_refd_to_none + grad_ontime_refd_to_none, 
                                    data = transition_dat)

# Graduation referenced to OnTime, w/ income and illiteracy
transition_simp_mod_m2_ontime <- lm( success ~ income + illiteracy + grad_late_refd_to_ontime + grad_none_refd_to_ontime, 
                                     data = transition_dat)
# Graduation referenced to None, w/ income and illiteracy
transition_simp_mod_m2_none  <- lm( success ~ income + illiteracy + grad_late_refd_to_none + grad_ontime_refd_to_none, 
                                    data = transition_dat)

# Model performance checks for all models
# performance::check_model(transition_simp_mod_m1_ontime)
# performance::check_model(transition_simp_mod_m1_none)
# performance::check_model(transition_simp_mod_m2_ontime)
# performance::check_model(transition_simp_mod_m2_none)

```


### Residuals & Assumptions

Descriptives section already checked normality so here we need to focus on linearity and homoscedasticity. We're  
going to center the data outright because I don't want to run everything multiple times.

```{r residuals_figures}

# Check Residuals plot (Heteroscedasticity)
# Fitted values vs. residuals to examine homoscedasticity
# NOTE use of .resid to plot residuals


# MODEL 1 RESIDUALS: transition ~ income + illiteracy + grad_refd_to_ontime
transition_resid_m1_ot_fig = ggplot( transition_simp_mod_m1_ontime, aes(.fitted, .resid) )

transition_resid_m1_ot_fig +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Transition Success Values (0 - 10)") +
    ylab("Success Residuals (0 - 10)") +
    ggtitle("Model 1 Transition Success Residual vs. Fitted Plot - Ref Ontime Grad") +
    my_gg_theme

# I also want to see a residuals plot in multiples of SD of productivity to help spot outliers
transition_resid_sd_m1_ot_fig = ggplot( transition_simp_mod_m1_ontime,
                                  aes(.fitted , .resid / transition_descr_comb$descriptives$asDF$`success[sd]` )
                                )

transition_resid_sd_m1_ot_fig +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Transition Success Values") +
    ylab(expression( "Residuals - Standardized (Multiples of " * sigma * ")") )  +
    ggtitle("Model 1 Transition Success Residual vs. Fitted Plot - Ref Ontime Grad") +
    my_gg_theme

# MODEL 1 RESIDUALS: transition ~ income + illiteracy + grad_refd_to_none
transition_resid_m1_no_fig = ggplot( transition_simp_mod_m1_none, aes(.fitted, .resid) )

transition_resid_m1_no_fig +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Transition Success Values (0 - 10)") +
    ylab("Success Residuals (0 - 10)") +
    ggtitle("Model 1 Transition Success Residual vs. Fitted Plot - Ref No Grad") +
    my_gg_theme

# I also want to see a residuals plot in multiples of SD of productivity to help spot outliers
transition_resid_sd_m1_no_fig = ggplot( transition_simp_mod_m1_none,
                                  aes(.fitted , .resid / transition_descr_comb$descriptives$asDF$`success[sd]` )
                                )

transition_resid_sd_m1_no_fig +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Transition Success Values") +
    ylab(expression( "Residuals - Standardized (Multiples of " * sigma * ")") )  +
    ggtitle("Model 1 Transition Success Residual vs. Fitted Plot - Ref No Grad") +
    my_gg_theme

# MODEL 2 RESIDUALS: transition ~ income + illiteracy + grad_refd_to_ontime
transition_resid_m2_ot_fig = ggplot( transition_simp_mod_m2_ontime, aes(.fitted, .resid) )

transition_resid_m2_ot_fig +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Transition Success Values (0 - 10)") +
    ylab("Success Residuals (0 - 10)") +
    ggtitle("Model 2 Transition Success Residual vs. Fitted Plot - Ref Ontime Grad") +
    my_gg_theme

# I also want to see a residuals plot in multiples of SD of productivity to help spot outliers
transition_resid_sd_m2_ot_fig = ggplot( transition_simp_mod_m2_ontime,
                                  aes(.fitted , .resid / transition_descr_comb$descriptives$asDF$`success[sd]` )
                                )

transition_resid_sd_m2_ot_fig +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Transition Success Values") +
    ylab(expression( "Residuals - Standardized (Multiples of " * sigma * ")") )  +
    ggtitle("Model 2 Transition Success Residual vs. Fitted Plot - Ref Ontime Grad") +
    my_gg_theme

# MODEL 2 RESIDUALS: transition ~ income + illiteracy + grad_refd_to_none
transition_resid_m2_no_fig = ggplot( transition_simp_mod_m2_none, aes(.fitted, .resid) )

transition_resid_m2_no_fig +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Transition Success Values (0 - 10)") +
    ylab("Success Residuals (0 - 10)") +
    ggtitle("Model 2 Transition Success Residual vs. Fitted Plot - Ref No Grad") +
    my_gg_theme

# I also want to see a residuals plot in multiples of SD of productivity to help spot outliers
transition_resid_sd_m2_no_fig = ggplot( transition_simp_mod_m2_none,
                                  aes(.fitted , .resid / transition_descr_comb$descriptives$asDF$`success[sd]` )
                                )

transition_resid_sd_m2_no_fig +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Transition Success Values") +
    ylab(expression( "Residuals - Standardized (Multiples of " * sigma * ")") )  +
    ggtitle("Model 2 Transition Success Residual vs. Fitted Plot - Ref No Grad") +
    my_gg_theme

```

### Homscedasticity Checks

Just to throw the kitchen sink at the problem, we'll run BP tests on all 4 models.

```{r homoscedasticity_checks}
# Run a Breusch Pagan Test of homoscedasticity for both models, for both reference grad values

cat( paste("  ", "---- Breusch-Pagan MODEL 1 - OnTime Ref ----", " ", sep = "\n") )
car::ncvTest(transition_simp_mod_m1_ontime)

cat( paste("  ", "---- Breusch-Pagan MODEL 1 - None Ref ----", " ", sep = "\n") )
car::ncvTest(transition_simp_mod_m1_none)

cat( paste("  ", "---- Breusch-Pagan MODEL 2 - OnTime Ref ----", " ", sep = "\n") )
car::ncvTest(transition_simp_mod_m2_ontime)

cat( paste("  ", "---- Breusch-Pagan MODEL 2 - None Ref ----", " ", sep = "\n") )
car::ncvTest(transition_simp_mod_m2_none)
```

### Data Cleaning - OPTIONAL
At first glance there was no need to revisit the data (clean the data), but this is kept here as a placeholder

``` {r data_cleaning_raw}

```

### Add Centered Data

We center our predictors for both datasets so regression tables can be built using centered data explicitly if so  
needed

```{r center_data}

# Center all continuous predictor data for models. We DON'T center outcome var (transition success) or categorical vars  
# (HS graduation times)
transition_dat$income_cent     <- transition_dat$income - mean(transition_dat$income)
transition_dat$illiteracy_cent <- transition_dat$illiteracy - mean(transition_dat$illiteracy)

```

### Build Multiple Regression Models

We're going to build up the regression model one predictor at a time, starting with HS graduation. We do technically    
have two "flavors" of model in how we're running this, as our primary model is referenced to graduating on time as the  
baseline. So we will create another section where we use the variables referenced against no graduation for comparison  

#### UnCentered - Referenced to OnTime

Uncentered data with OnTime graduation as reference categorical value

```{r multiple_regression_uncentered_ref_ot}
cat( 
     paste(
           '--------------------------------------------------------------------------------',
           '---- UnCentered Multiple Regression Model w/ OnTime Graduation as Reference ----',
           '--------------------------------------------------------------------------------',
           sep = '\n'
     ) 
   )

# OnTime Graduation as Reference. UnCentered.   
tranistion_mult_regr_ot <- linReg( transition_dat,
                                   dep = 'success',
                                   covs = c('income','illiteracy', 'grad_late_refd_to_ontime', 'grad_none_refd_to_ontime'),
                                   blocks = list(
                                                  list( 'grad_late_refd_to_ontime', 'grad_none_refd_to_ontime' ),
                                                  list( 'income' ),
                                                  list( 'illiteracy' )
                                                 ),
                                   modelTest = TRUE,
                                   stdEst = TRUE,
                                   r2Adj = TRUE,
                                   collin = TRUE,
                                   ci = TRUE)
tranistion_mult_regr_ot

```

#### Centered - Referenced to OnTime

Centered data with OnTime graduation as reference categorical value

```{r multiple_regression_centered_ref_ot}
cat( 
     paste(
           '------------------------------------------------------------------------------',
           '---- Centered Multiple Regression Model w/ OnTime Graduation as Reference ----',
           '------------------------------------------------------------------------------',
           sep = '\n'
     ) 
   )

# OnTime Graduation as Reference. Centered.        
tranistion_mult_regr_ot_cent <- linReg( transition_dat,
                                        dep = 'success',
                                        covs = c('income_cent','illiteracy_cent', 'grad_late_refd_to_ontime', 'grad_none_refd_to_ontime'),
                                        blocks = list(
                                                       list( 'grad_late_refd_to_ontime', 'grad_none_refd_to_ontime' ),
                                                       list( 'income_cent' ),
                                                       list( 'illiteracy_cent' )
                                                      ),
                                        modelTest = TRUE,
                                        stdEst = TRUE,
                                        r2Adj = TRUE,
                                        collin = TRUE,
                                        ci = TRUE)
tranistion_mult_regr_ot_cent

```


#### UnCentered - Referenced to No Graduation

Uncentered data with NO graduation as reference categorical value

```{r multiple_regression_uncentered_ref_no}
cat( 
     paste(
           '--------------------------------------------------------------------------------',
           '---- UnCentered Multiple Regression Model w/ None Graduation as Reference ----',
           '--------------------------------------------------------------------------------',
           sep = '\n'
     ) 
   )

# OnTime Graduation as Reference. UnCentered.   
tranistion_mult_regr_no <- linReg( transition_dat,
                                   dep = 'success',
                                   covs = c('income','illiteracy', 'grad_late_refd_to_none', 'grad_ontime_refd_to_none'),
                                   blocks = list(
                                                  list( 'grad_late_refd_to_none', 'grad_ontime_refd_to_none' ),
                                                  list( 'income' ),
                                                  list( 'illiteracy' )
                                                 ),
                                   modelTest = TRUE,
                                   stdEst = TRUE,
                                   r2Adj = TRUE,
                                   collin = TRUE,
                                   ci = TRUE)
tranistion_mult_regr_no

```

#### Centered - Referenced to No Graduation

Centered data with NO graduation as reference categorical value

```{r multiple_regression_centered_ref_no}
cat( 
     paste(
           '------------------------------------------------------------------------------',
           '---- Centered Multiple Regression Model w/ None Graduation as Reference ----',
           '------------------------------------------------------------------------------',
           sep = '\n'
     ) 
   )

# OnTime Graduation as Reference. Centered.        
tranistion_mult_regr_no_cent <- linReg( transition_dat,
                                        dep = 'success',
                                        covs = c('income_cent','illiteracy_cent', 'grad_late_refd_to_none', 'grad_ontime_refd_to_none'),
                                        blocks = list(
                                                       list( 'grad_late_refd_to_none', 'grad_ontime_refd_to_none' ),
                                                       list( 'income_cent' ),
                                                       list( 'illiteracy_cent' )
                                                      ),
                                        modelTest = TRUE,
                                        stdEst = TRUE,
                                        r2Adj = TRUE,
                                        collin = TRUE,
                                        ci = TRUE)
tranistion_mult_regr_no_cent

```

### APA Tables

Output regression tables to an APA formatted (almost) word doc in case that's useful for all models

```{r apa_tables}
# Our simple linear models are still defined from the prior section so we can just plug them in here.
# Use reg.table method in APA package to generate regression table

# Document tables for success ~ income + grad
apa.reg.table( transition_simp_mod_m1_ontime, filename="transition_inc_grad_ref_ontime.doc", table.number = 2 )
apa.reg.table( transition_simp_mod_m1_none,   filename="transition_inc_grad_ref_none.doc", table.number = 2 )

# Document tables for success ~ income + grad
apa.reg.table( transition_simp_mod_m2_ontime, filename="transition_inc_ill_grad_ref_ontime.doc", table.number = 2 )
apa.reg.table( transition_simp_mod_m2_none,   filename="transition_inc_ill_grad_ref_none.doc", table.number = 2 )
```
