---
title: "PSYCH308D - Data Analysis 01 (DA01)"
author: "Brady C. Jackson"
date: "2025/03/27"

# Write document output to HTML, Word, and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Prompt
Parents with anxiety disorders tend to have children with anxiety disorders as well (see everything ever written by  
Bogels, Borelli, Wood, or Rapee). This has led some to hypothesize that anxiety is genetically transmitted. However,  
you think there might be something else going on. Parental overcontrol, the excess regulation of a child's emotion,  
cognition, and behavior is strongly related to child anxiety. You wonder if anxious parents are more overcontrolling,  
leading to child anxiety. You decided to conduct a study to figure this out. 
 
## Variables 

Input Variable Table HERE

| Variable | Type | Description |
|:-----|:---|:------------------|
| `Parent_Anx` | 1-20 scale | PV: Higher scores indicating higher anxiety symptoms      |
| `Child_Anx`  | 1-20 scale | OV: Higher scores indicating higher anxiety symptoms      |
| `Parent_OC`  | 1-20 scale | MedV: Higher scores indicating greater use of overcontrol |

## Research Question  
  
Does parental overcontrol mediate the relationship between parent anxiety and child anxiety?

## Assignment

### Part 1

Conduct the appropriate analysis using *both* the *Baron & Kenney method* and *bootstrapping techniques* and write an    
APA results section. Report all the proper assumptions and statistics in the results section. Don't forget to include    
all appropriate tables and figures for both methods. 

### Part 2

*Respond to the following questions.*

1. Interpret your findings for a lay audience. 
2. Run Sobels test and compare your results to your bootstrapped analysis findings.   
   If the two analyses found different outcomes, determine which results are more appropriate and justify your   
   decision. (Hint: What is a known issue with the Sobel test?)
3. If you bootstrapped a mediation analysis with 1000 iterations twice, would you obtain the same exact statistics  
   both times? Explain why or why not. 

---ONLY WRITE CODE BELOW THIS LINE---

## Code

### Libraries

Load all requisite libraries here.

```{r package_loading, message=FALSE, warning=FALSE}
# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(tidyverse)
library(jmv)       # for descriptive
library(ggplot2)
library(dplyr)
library(psych)
library(corrplot)    # For fancy covariance matrix plots
library(car)         # for ncvTest (Breusch Pagan)
library(stringr)     # for sub_str operations
library(Hmisc)       # for fun.dat substitution
library(see)         # for outliers analysis 
library(magrittr)
library(foreign)
library(broom)
library(robmed)
library(mediation)   # For mediation analysis
library(multilevel)
library(GGally)
library(lsr)
library(car)
library(mvnTest)     # Multivariate Normality
library(lm.beta)     # For extracting coefficients from bootstrap runs

```

### Metadata

This section of code is to setup some general variables that we'll use throughout the code (e.g. figure colors, etc)

```{r metadata}
# First we'll defines some meta-data to use in all of our plots so they're nice and clean
font_color = "#4F81BD"
grid_color_major = "#9BB7D9"
grid_color_minor = "#C8D7EA"
back_color = "gray95"
rb_colmap = colorRampPalette( c("firebrick", "grey86", "dodgerblue3") )(200)

# I'm going to try to save off my preferred ggplot theme combinations as a unqiue theme object that I can just reference
# later in the code....totally unclear if ggplot works this way....
my_gg_theme = theme_minimal() +
              theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
                     axis.title.x = element_text(color = font_color),
                     axis.title.y = element_text(color = font_color),
                     axis.text.x = element_text(color = font_color),
                     axis.text.y = element_text(color = font_color),
                     legend.title = element_text(color = font_color),
                     legend.text = element_text(color = font_color),
                     panel.grid.minor = element_line(color = grid_color_minor),
                     panel.grid.major = element_line(color = grid_color_major),
                     panel.background = element_rect(fill = back_color, color = font_color)
                   )

```

### Load and View data

Here we load the dataframe from the CSV file provided for the assignment, and we 


```{r load_raw_data}

# Load the assignment data from CSV
raw_dat = read.csv("./308D.DA1 Mediation.Data.csv")

# Rename columns to lower because why not
colnames(raw_dat) <- tolower( colnames(raw_dat) )

# View data
# glimpse(breaks_raw_dat)

# Ensure that the numbers of each subject in the study are unique to prevent any duplicate data
# if the size of the unique-entries only is the same as the whole vector then there are no duplicate subjects
# NOTE: This fails if the colname of the subject ID is input wrong (e.g. if the subject ID is not named "participant"
#       in the dataframe. So check that the colname exists before blindly believing)
test_unique = (length(unique(raw_dat$participant)) == length(raw_dat$participant))
if(!test_unique){
    print("WARNING: There are duplicate data entries in the raw data")
}else{
    print("No duplicate entries detected in raw data")
}
```

### Add Centered Data

It is often useful to have the centered data in addition to the uncentered data. So we add that here manually.

```{r center_data}

# Center all continuous data, which, in our dataframe, is the predictor variable, outcome variable, and mediator variable
raw_dat$parent_anx_cent <- raw_dat$parent_anx - mean(raw_dat$parent_anx)
raw_dat$parent_oc_cent  <- raw_dat$parent_oc - mean(raw_dat$parent_oc)
raw_dat$child_anx_cent  <- raw_dat$child_anx - mean(raw_dat$child_anx)
```

### Construct Standardized Dataframe and add Centered Values

For the purposes of mediation models, we need manually defined models that rely on standardized data (centered and   
scaled by standard deviation). We could compute this manually after extracting the mean and SD from the descriptives   
dataframe, but the "scale" function builds this capability automagically in R so we just use it here.

```{r standardize_variables}

# We can use the scale function to center and standardize the data in the original dataframe
# NOTE: We have to access the contents of the dataframe returned by scale with [,1] as scale will natrually return a 
#       whole dataframe, not just a numeric array. So this gets everything defined as the correct datatype to match the  
#       other contents in raw_dat
raw_dat$parent_anx_std = scale(raw_dat$parent_anx, center = TRUE, scale = TRUE)[,1]
raw_dat$parent_oc_std = scale(raw_dat$parent_oc, center = TRUE, scale = TRUE)[,1]
raw_dat$child_anx_std = scale(raw_dat$child_anx, center = TRUE, scale = TRUE)[,1]
```

### Descriptive Statistics - Raw Data

This section will look at base descriptive statistics of the raw data to help identify data anomalies and check     
normality of variables.   
  NOTE: While including the descriptives of the centered and standardized vars looks redundant, it gives us a way of   
        confirming those computations did what we thought they would.

```{r descriptive_stats_raw}

# We'll need to check both univariate normality and multi-variate normality.
my_descr = jmv::descriptives( raw_dat,
                              vars = c('parent_anx', 'parent_oc', 'child_anx', 
                                       'parent_anx_cent', 'parent_oc_cent', 'child_anx_cent', 
                                       'parent_anx_std', 'parent_oc_std', 'child_anx_std'),
                              hist = TRUE,
                              dens = TRUE,
                              qq = TRUE,
                              sd = TRUE,
                              variance = TRUE,
                              se = TRUE,
                              skew = TRUE,
                              kurt = TRUE,
                              missing = TRUE
                            )
print(my_descr)

```

### Correlation Plots - Raw Data

Visualize the covariance matrix to understand correlations between all numeric variables. Note we have to manually  
remove categorical variables from this output.


```{r correlation_plots_raw}

# We look at correlations for both the raw data as they will be equivalent between both centered and uncenterd
# variations
raw_subset = raw_dat[ c('parent_anx', 'parent_oc', 'child_anx') ]
corr_raw <- stats::cor( raw_subset )
corrplot( corr_raw,
          method="color",
          type = "full",
          addCoef.col = "black",
          col = rb_colmap,
          tl.col = font_color )

# Print correlation tables (annoying that we have to use two different correlation functions, 1 for plotting and
# 1 for tables, but oh well)
cor_tab_raw <- jmv::corrMatrix(raw_subset, flag = TRUE)
print( cor_tab_raw )

```

### Scatterplots of Variables

Since we're basing our models of assumed linear relationships we need to look at scatterplots of our various PV->OV  
relationships. In  making mediation models, we'll be building the following 3x regressions:
 OV   ~ PV   
 OV   ~ PV + MedV   
 MedV ~ PV   
   
 So we need to build each of these scatterplots

```{r scatter_plots}

# All scatterplots are uncentered 

# Scatterplot 1: Child Anxiety vs. Parent Anxiety  (uncentered)
scatter_ca_v_pa <- ggplot(raw_dat, aes(child_anx, parent_anx) )

scatter_ca_v_pa +
    geom_point() +
    geom_smooth(method = "lm", colour = "Red") +
    ggtitle("Childhood Anxiety as a Function of Parent Anxiety") +
    labs(y = "Childhood Anxiety (1-20)", x = "Parent Anxiety (1-20)") +
    my_gg_theme


# Scatterplot 2: Child Anxiety vs. Parent Overcontrol  (uncentered)
scatter_ca_v_po <- ggplot(raw_dat, aes(child_anx, parent_oc) )

scatter_ca_v_po +
    geom_point() +
    geom_smooth(method = "lm", colour = "Red") +
    ggtitle("Childhood Anxiety as a Function of Parent Overcontrol") +
    labs(y = "Childhood Anxiety (1-20)", x = "Parent Overcontrol (1-20)") +
    my_gg_theme

# Scatterplot 3: Parent Overcontrol vs. Parent Anxiety  (uncentered)
scatter_po_v_pa <- ggplot(raw_dat, aes(parent_oc, parent_anx) )

scatter_po_v_pa +
    geom_point() +
    geom_smooth(method = "lm", colour = "Red") +
    ggtitle("Parent Overcontrol as a Function of Parent Anxiety") +
    labs(y = "Parent Overcontrol (1-20)", x = "Parent Anxiety (1-20)") +
    my_gg_theme

# Make sure you plot OV as fncn of both PV and Mediator.

# Make sure you plot Mediator as function of PV

```

### Build Simple Linear Models (lm)

For the purpose of B&K, bootstrapping, and assumption checks we need to define multiple linear models. Some will be
in the unstandardized space, while others are in the standardized space

```{r linear_model_definition}

# # We create a simple linear model using the two main effect predictors (FA and PPR) w/o accounting for moderation
# # to aid in simple data assumption checks later on (homoscedasticity and multi-variate normality.)
#  
# # Aggression regressed onto family adversity and positive peer support as main effect predictors.
# agg_simp_mod_uc <- lm( agg ~ fa + ppr, data = this_dat)
# 
# # Model performance checks for all models
# performance::check_model(agg_simp_mod_uc)
# 
# # We'll create a simple model that includes the moderation term as well
# # NOTE: We're using uncentered data here:
# agg_me_w_mod_model_uc <- lm(agg ~ fa + ppr + p_mod_f, data = this_dat)
# # Notice funny grouping when we define model as above. Spot checked these alternative models and it's not clear which is  
# # better.
# # agg_me_w_mod_model_uc <- lm(agg ~ p_mod_f, data = this_dat)
# # agg_me_w_mod_model_uc <- lm(agg ~ fa_cent + ppr_cent + p_mod_f_cent, data = this_dat)

# Unstandardized linear model definition

# Unstandardized Path a
mod_po_v_pa_unstd = lm(parent_oc ~ parent_anx, data = raw_dat)

# Unstandardized Paths b + c'
mod_ca_v_pa_and_po_unstd = lm(child_anx ~ parent_anx + parent_oc, data = raw_dat)

# Standardized linear model definition

# Standardized Path a
mod_po_v_pa_std = lm(parent_oc_std ~ parent_anx_std, data = raw_dat)

# Standardized Path b + c'
mod_ca_v_pa_and_po_std = lm(child_anx_std ~ parent_anx_std + parent_oc_std, data = raw_dat)
```

### Residuals Plots

For the purposes of heteroscedasticity, we're going to look at unstandardized, saturated models. The saturated model  
in mediation models are the b + c' path.

```{r residuals_figures}

# Check Residuals plot (Heteroscedasticity)
# Fitted values vs. residuals to examine homoscedasticity
# NOTE use of .resid to plot residuals


# Unstandardized Residuals Plot: Childhood anxiety
fig_ca_v_pa_and_po_unstd = ggplot( mod_ca_v_pa_and_po_unstd, aes(.fitted, .resid) )

fig_ca_v_pa_and_po_unstd +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Childhood Anxiety Levels (1-20)") +
    ylab("Childhood Anxiety Residuals (1 - 20)") +
    ggtitle("Unstandardized - Childhood Anxiety Residual vs. Fitted Plot") +
    my_gg_theme

# I also want to see a residuals plot in multiples of SD of productivity to help spot outliers
# We can use our standardized models here to check this.
fig_ca_v_pa_and_po_std = ggplot( mod_ca_v_pa_and_po_std, aes(.fitted, .resid) )

fig_ca_v_pa_and_po_std +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Childhood Anxiety Levels (Multiples of SD)") +
    ylab("Childhood Anxiety Residuals (Multiples of SD)") +
    ggtitle("Standardized - Childhood Anxiety Residual vs. Fitted Plot") +
    my_gg_theme

# Finally, we need to check the same residuals plots for parental overcontrol as a function of parental anxiety 
# (Mediator regressed onto PV)

# Unstandardized Residuals Plot: Parental Overcontrol
fig_po_v_pa_unstd = ggplot( mod_po_v_pa_unstd, aes(.fitted, .resid) )

fig_po_v_pa_unstd +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Parental Overcontrol Levels (1-20)") +
    ylab("Parental Overcontrol Residuals (1 - 20)") +
    ggtitle("Unstandardized - Parental Overcontrol Residual vs. Fitted Plot") +
    my_gg_theme

# Standardized Residuals Plot: Parental Overcontrol
fig_po_v_pa_std = ggplot( mod_po_v_pa_std, aes(.fitted, .resid) )

fig_po_v_pa_std +
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Parental Overcontrol Levels (Multiples of SD)") +
    ylab("Parental Overcontrol Residuals (Multiples of SD)") +
    ggtitle("Standardized - Parental Overcontrol Residual vs. Fitted Plot") +
    my_gg_theme

```


### Homoscedasticity Checks
```{r homoscedasticity_checks}

# Run a Breusch Pagan Test of homoscedasticity for both the saturated model, and the mediator model
#  NOTE: Results will be the same for equivalent standardized and unstandardized models
cat( paste("  ", "---- Breusch-Pagan Path b + c' ----", " ", sep = "\n") )
car::ncvTest(mod_ca_v_pa_and_po_unstd)

cat( paste("  ", "---- Breusch-Pagan Path a ----", " ", sep = "\n") )
car::ncvTest(mod_po_v_pa_unstd)
```


### Multi-Variate Normality

```{r multivar_norm}

# Code below manually produces QQ Plot for better alignment with general plot styles
# Unstandardized QQ plot for saturated (path b + c', model)
ggplot(mod_ca_v_pa_and_po_unstd,
         aes(
               qqnorm( .stdresid,
                       xlab = "", ylab = "",
                       conf.level = 0.945,
                       conf.args = list(col = "lightgrey", exact = TRUE)
                      )[[1]],
               .stdresid )
       ) +
    geom_point(na.rm = TRUE) +
    xlab("Theoretical Quantiles") +
    ylab("Standardized Residuals") +
    ggtitle("Normal Q-Q: Childhood Anxiety ~ Parental Anxiety + Overcontrol") +
    my_gg_theme

# Unstandardized QQ plot for saturated (path a, model)
ggplot(mod_po_v_pa_unstd,
         aes(
               qqnorm( .stdresid,
                       xlab = "", ylab = "",
                       conf.level = 0.945,
                       conf.args = list(col = "lightgrey", exact = TRUE)
                      )[[1]],
               .stdresid )
       ) +
    geom_point(na.rm = TRUE) +
    xlab("Theoretical Quantiles") +
    ylab("Standardized Residuals") +
    ggtitle("Normal Q-Q: Parental Overcontrol ~ Parental Anxiety") +
    my_gg_theme

# Henze-Zirkler test of multivariate normality
# Columns 2-4 are Parental Anxiety, Parental Overcontrol, and Childhood Anxiety - all unstandardized
HZ.test(raw_dat[2:4])


```

### Data Cleaning - OPTIONAL

After reviewing all assumptions and data homogeneity matters above, found no need to dataclean

``` {r data_cleaning_raw}

```

### Construct Mediation Models

We need to perform multiple mediation analysis methods, including B&K's OG method, that same supplemented with Sobel's  
test, and bootstrapping our best answer. For bootstrapping, we'll want both standardized and unstandardized coefficient  
outputs which means we'll need to define our linear regression models with both standardized data and unstandardized   
data.

#### Baron and Kenny - Unstandardized Data

Here we use B&K's method to evaluate the impact of media


```{r baron_and_kenny_method}

# SAMPLE CODE - BARON and KENNY

# #Step 1: Regress X (treat_num) onto Y (depress2)
# model.j1 <- linReg(data = dat.jobs, dep = 'depress2', 
#                  covs = c('treat_num', 'depress1'), 
#                  blocks = list(
#                    list('treat_num', 'depress1')), tion
#                 collin = TRUE, 
#                 stdEst = TRUE,
#                 ci = TRUE)
# model.j1
# 
# #Step 2: Regress M (mastery) onto X (treat_num)
# model.j2 <- linReg(data = dat.jobs, dep = 'mastery', 
#                  covs = c('treat_num'), 
#                  blocks = list(
#                    list('treat_num')), 
#                 collin = TRUE, 
#                 stdEst = TRUE,
#                 ci = TRUE)
# model.j2
# 
# #Step 3: Regress X (treat_num) and M (mastery) onto Y (depress2)
# model.j3 <- linReg(data = dat.jobs, dep = 'depress2', 
#                  covs = c('treat_num', 'mastery', 'depress1'), 
#                  blocks = list(
#                    list('treat_num', 'mastery', 'depress1')), 
#                 collin = TRUE, 
#                 stdEst = TRUE,
#                 ci = TRUE)
# model.j3

# Model 1: Evaluate Total Effect (Path c), regress OV onto PV
regr_ca_v_pa <- linReg( raw_dat,
                        dep = 'child_anx',
                        covs = c('parent_anx'),
                        blocks = list(
                                       list( 'parent_anx' )
                                      ),
                        modelTest = TRUE,
                        stdEst = TRUE,
                        r2Adj = TRUE,
                        collin = TRUE,
                        ci = TRUE)
regr_ca_v_pa

# Model 2: Evaluate Path a (regress mediator onto PV)
regr_po_v_pa <- linReg( raw_dat,
                        dep = 'parent_oc',
                        covs = c('parent_anx'),
                        blocks = list(
                                       list( 'parent_anx' )
                                      ),
                        modelTest = TRUE,
                        stdEst = TRUE,
                        r2Adj = TRUE,
                        collin = TRUE,
                        ci = TRUE)
regr_po_v_pa

# Model 3: Evaluate Path b and c'  regress OV onto Mediator and PV. We build up two models here, recasting Model 1 
#          above, so we can directly compare it.
regr_ca_v_pa_and_po <- linReg( raw_dat,
                               dep = 'child_anx',
                               covs = c('parent_anx', 'parent_oc'),
                               blocks = list(
                                              list( 'parent_anx' ),
                                              list( 'parent_oc' )
                                             ),
                               modelTest = TRUE,
                               stdEst = TRUE,
                               r2Adj = TRUE,
                               collin = TRUE,
                               ci = TRUE)
regr_ca_v_pa_and_po

```

#### Sobel's Test

Now we need to perform the Null Hypothesis test to evaluate if the a+b effects are significantly discernable from 0  

```{r sobel_test}

# Test the indirect path with sobels test based on previous section models
#  NOTE: We don't need to define a covariate here since we don't have any control variables.
#        Additionally, we use the m(MedV) syntax to define the mediator per the help docs       
test_mediation( child_anx ~ parent_anx + m(parent_oc),
                test = c("sobel"), 
                data = raw_dat
              )


```

#### Bootstrap's Bootstraps - Unstandardized Bootstraps

Our standardized and unstandardized models are already defined, so here we can focus on the boostrapping code

```{r unstandardized_bootstrap}
# Define number of sims to run
nsims = 10000

# Unstandardized Bootstrap
# Estimate parameters of the model using a bootstrap sampling method.
#  Our mediator linear model is passed first (a), followed by our outcome variable saturated model (b + c')
#  Asking for 10,0000 bootstrap simulations with confidence intervals ("perc"). CI level set at 95%
#  Need to tell R which variable is the predictor and which is the mediator ("mediator"). 
#  Adding in identity centrality as a covariate so we can control for this ("treat").
bootstrap_unstd <- mediation::mediate( mod_po_v_pa_unstd, mod_ca_v_pa_and_po_unstd,
                                       treat = "parent_anx", 
                                       mediator = "parent_oc", 
                                       sims = nsims, 
                                       boot = TRUE, 
                                       boot.ci.type = "perc",
                                       conf.level = 0.95, 
                                       long = TRUE
                                     )

#summarize the parameter estimates
summary(bootstrap_unstd)

# This produces the individual a and b path betas
unstd_coeff_a = lm.beta(mod_po_v_pa_unstd)
unstd_coeff_b = lm.beta(mod_ca_v_pa_and_po_unstd)

cat( paste( "  ", "---- Unstandardized Coefficients a and b ----", " ", 
            "Path a Coefficient (B_a): ", unstd_coeff_a$coefficients["parent_anx"],
            "Path b Coefficient (B_b): ", unstd_coeff_b$coefficients["parent_oc"],
            "   ",
            sep = "\n"
          ) 
   )




```



#### Bootstrap me Baby, One More Time - Standardized Bootstraps

Our standardized and unstandardized models are already defined, so here we can focus on the boostrapping code

```{r standardized_bootstrap}
# Standardized Bootstrap
# Estimate parameters of the model using a bootstrap sampling method.
#  Our mediator linear model is passed first (a), followed by our outcome variable saturated model (b + c')
#  Asking for 10,0000 bootstrap simulations with confidence intervals ("perc"). CI level set at 95%
#  Need to tell R which variable is the predictor and which is the mediator ("mediator"). 
#  Adding in identity centrality as a covariate so we can control for this ("treat").
bootstrap_std <- mediation::mediate( mod_po_v_pa_std, mod_ca_v_pa_and_po_std,
                                     treat = "parent_anx_std", 
                                     mediator = "parent_oc_std", 
                                     sims = nsims, 
                                     boot = TRUE, 
                                     boot.ci.type = "perc",
                                     conf.level = 0.95, 
                                     long = TRUE
                                   )

#summarize the parameter estimates
summary(bootstrap_std)

# This produces the individual a and b path betas
std_coeff_a = lm.beta(mod_po_v_pa_std)
std_coeff_b = lm.beta(mod_ca_v_pa_and_po_std)

cat( paste( "  ", "---- Standardized Coefficients a and b ----", " ", 
            "Path a Coefficient (Beta_a): ", std_coeff_a$standardized.coefficients["parent_anx_std"],
            "Path b Coefficient (Beta_b): ", std_coeff_b$standardized.coefficients["parent_oc_std"],
            "   ",
            sep = "\n"
          ) 
   )

```






