---
title: "PSYCH308D - Data Analysis 01 (DA01)"
author: "Brady C. Jackson"
date: "2025/03/27"

# Write document output to HTML, Word, and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Prompt
Parents with anxiety disorders tend to have children with anxiety disorders as well (see everything ever written by  
Bogels, Borelli, Wood, or Rapee). This has led some to hypothesize that anxiety is genetically transmitted. However,  
you think there might be something else going on. Parental overcontrol, the excess regulation of a child's emotion,  
cognition, and behavior is strongly related to child anxiety. You wonder if anxious parents are more overcontrolling,  
leading to child anxiety. You decided to conduct a study to figure this out. 
 
## Variables 

Input Variable Table HERE

| Variable | Type | Description |
|:-----|:---|:------------------|
| `Parent_Anx` | 1-20 scale | Higher scores indicating higher anxiety symptoms |
| `Child_Anx`  | 1-20 scale | Higher scores indicating higher anxiety symptoms    |
| `Parent_OC`  | 1-20 scale | Higher scores indicating greater use of overcontrol |

## Research Question  
  
Does parental overcontrol mediate the relationship between parent anxiety and child anxiety?

## Assignment

### Part 1

Conduct the appropriate analysis using *both* the *Baron & Kenney method* and *bootstrapping techniques* and write an    
APA results section. Report all the proper assumptions and statistics in the results section. Don't forget to include    
all appropriate tables and figures for both methods. 

### Part 2

*Respond to the following questions.*

1. Interpret your findings for a lay audience. 
2. Run Sobels test and compare your results to your bootstrapped analysis findings.   
   If the two analyses found different outcomes, determine which results are more appropriate and justify your   
   decision. (Hint: What is a known issue with the Sobel test?)
3. If you bootstrapped a mediation analysis with 1000 iterations twice, would you obtain the same exact statistics  
   both times? Explain why or why not. 

---ONLY WRITE CODE BELOW THIS LINE---

## Code

### Libraries

Load all requisite libraries here.

```{r package_loading, message=FALSE, warning=FALSE}
# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(tidyverse)
library(jmv)       # for descriptive
library(ggplot2)
library(dplyr)
library(psych)
library(corrplot)    # For fancy covariance matrix plots
library(car)         # for ncvTest (Breusch Pagan)
library(stringr)     # for sub_str operations
library(Hmisc)       # for fun.dat substitution
library(see)         # for outliers analysis 
library(magrittr)
library(foreign)
library(broom)
library(robmed)
library(mediation)   # For mediation analysis
library(multilevel)
library(GGally)
library(lsr)
library(car)
library(mvnTest)     # Multivariate Normality
library(lm.beta)

```

### Metadata

This section of code is to setup some general variables that we'll use throughout the code (e.g. figure colors, etc)

```{r metadata}
# First we'll defines some meta-data to use in all of our plots so they're nice and clean
font_color = "#4F81BD"
grid_color_major = "#9BB7D9"
grid_color_minor = "#C8D7EA"
back_color = "gray95"
rb_colmap = colorRampPalette( c("firebrick", "grey86", "dodgerblue3") )(200)


```

### Load and View data

```{r load_raw_data}

# SAMPLE CODE
# # Stuff goes here
# breaks_raw_dat = read.csv("./da01.csv")
# 
# # Rename columns to lower because why not
# colnames(breaks_raw_dat) <- tolower( colnames(breaks_raw_dat) )
# 
# # View data
# glimpse(breaks_raw_dat)
# 
# # Ensure that the employee numbers of each subject in the study are unique to prevent any duplicate data
# # if the size of the unique-entries only is the same as the whole vector then there are no duplicate subjects
# test_unique = (length(unique(breaks_raw_dat$subject)) == length(breaks_raw_dat$subject))
# if(!test_unique){
#     print("WARNING: There are duplicate data entries in the raw data")
# }else{
#     print("No duplicate entries detected in raw data")
# }
```

### Descriptive Statistics - Raw Data

This section will look at base descriptive statistics of the raw data to help identify data anomalies and check   
normality of predictor variables (break length - minutes)

```{r descriptive_stats_raw}

# SAMPLE CODE
# # Get JMV descriptives of raw productivity, enjoyment, and break length
# # Tabulate mean, variance, skew, and kurtosis so we can check normality.
# # Output histograms and qq plots so we can spot outliers and review distribution shape
# breaks_descr = jmv::descriptives( breaks_raw_dat[c(2:4)],
#                                   hist = TRUE,
#                                   dens = TRUE,
#                                   qq = TRUE,
#                                   sd = TRUE, 
#                                   variance = TRUE,
#                                   se = TRUE, 
#                                   skew = TRUE, 
#                                   kurt = TRUE,
#                                   missing = TRUE)
# print(breaks_descr)

```

### Correlation Plots - Raw Data

Visualize the covariance matrix to understand correlation between break length and employee productivity


```{r correlation_plots_raw}

# SAMPLE CODE
# # Given that we're working explicitly on break length and productivity, I want to create a second dataframe that 
# # consists only of those columns for simplicity. I'll maintain the original as well so I can get some insight into
# # employee enjoyment as well, but since that's not the focus I want a clean dataset as well.
# breaks_dat <- breaks_raw_dat
# breaks_dat$subject <- NULL
# breaks_dat$enjoy <- NULL
# 
# # Produce a correlation matrix plot that includes the numbers printed with a color gradient for easy reading.
# # par(bg = "lightgrey")
# breaks_full_cor <- stats::cor(breaks_raw_dat)
# corrplot( breaks_full_cor, 
#           method="color", 
#           type = "full", 
#           addCoef.col = "black", 
#           col = rb_colmap, 
#           tl.col = font_color )
# 
# breaks_short_cor <- stats::cor(breaks_dat)
# corrplot( breaks_short_cor, 
#           method="color", 
#           type = "full",  
#           addCoef.col = "black", 
#           col = rb_colmap,  
#           tl.col = font_color )
# 
# 
# # Print correlation tables (annoying that we have to use two different correlation functions, 1 for plotting and
# # 1 for tables, but oh well)
# breaks_full_cor_tab <- jmv::corrMatrix(breaks_raw_dat[2:4], flag = TRUE)
# print(breaks_full_cor_tab)
# 
# breaks_short_cor_tab <- jmv::corrMatrix(breaks_dat, flag = TRUE)
# print(breaks_short_cor_tab)
# 
# # Finally dump a correlation table to Word Doc format for easy table formatting / to save time in writing
# # the assignment
# apa.cor.table( breaks_raw_dat[2:4], 
#                filename = "full_corr_table_breaks_v_prod.doc", 
#                table.number = 1, 
#                show.sig.stars = TRUE, 
#                landscape = TRUE)
# 
# apa.cor.table( breaks_dat, 
#                filename = "short_corr_table_breaks_v_prod.doc", 
#                table.number = 1, 
#                show.sig.stars = TRUE, 
#                landscape = TRUE)

```

### Assumptions, Figures, and Plots - Raw Data

Descriptives section already checked normality so here we need to focus on linearity and homoscedasticity. We're  
going to center the data outright because I don't want to run everything multiple times.

```{r assumptions_and_figures_raw}

# SAMPLE CODE
# # Centering the data
# breaks_dat$length_centered <- breaks_dat$length - mean(breaks_dat$length)
# 
# 
# # Check Scatterplot of Centered data: Break Length is PV, Productivity is OV
# breaks_raw_scatter <- ggplot(breaks_dat, aes(length, product) )
# 
# breaks_raw_scatter + 
#     geom_point() + 
#     geom_smooth(method = "lm", colour = "Red") + 
#     ggtitle("Break Length Relation to Employee Productivity") + 
#     labs(x = "Break Length (m)", y = "Employee Productivity (0-100)") +
#     theme_minimal() +
#     theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
#            axis.title.x = element_text(color = font_color),
#            axis.title.y = element_text(color = font_color),
#            axis.text.x = element_text(color = font_color),
#            axis.text.y = element_text(color = font_color),
#            legend.title = element_text(color = font_color),
#            legend.text = element_text(color = font_color),
#            panel.grid.minor = element_line(color = grid_color_minor),
#            panel.grid.major = element_line(color = grid_color_major),
#            panel.background = element_rect(fill = back_color, color = font_color)
#          )
# 
# breaks_cent_scatter <- ggplot(breaks_dat, aes(length_centered, product) )
# 
# breaks_cent_scatter + 
#     geom_point() + 
#     geom_smooth(method = "lm", colour = "Red") + 
#     ggtitle("Break Length Relation to Employee Productivity") + 
#     labs(x = "Break Length - Centered (m)", y = "Employee Productivity (0-100)") +
#     theme_minimal() +
#     theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
#            axis.title.x = element_text(color = font_color),
#            axis.title.y = element_text(color = font_color),
#            axis.text.x = element_text(color = font_color),
#            axis.text.y = element_text(color = font_color),
#            legend.title = element_text(color = font_color),
#            legend.text = element_text(color = font_color),
#            panel.grid.minor = element_line(color = grid_color_minor),
#            panel.grid.major = element_line(color = grid_color_major),
#            panel.background = element_rect(fill = back_color, color = font_color)
#          )
# 
# 
# # Check Residuals plot (Heteroscedasticity)
# # To compute residuals we need to define the model as productivity regressed onto break length in a linear fashion
# breaks_model_lin <- lm(product ~ length, data = breaks_dat)
# 
# # Fitted values vs. residuals to examine homoscedasticity
# # NOTE use of .resid to plot residuals
# breaks_resid_fig = ggplot( breaks_model_lin, aes(.fitted, .resid) )
# 
# breaks_resid_fig + 
#     geom_point(col = font_color) +
#     geom_hline(yintercept=0, col="green3", linetype="dashed") +
#     xlab("Fitted Productivity Values (0 - 100)") +
#     ylab("Productivity Residuals (0 - 100)") +
#     ggtitle("Residual vs. Fitted Plot") +
#     theme_minimal() +
#     theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
#            axis.title.x = element_text(color = font_color),
#            axis.title.y = element_text(color = font_color),
#            axis.text.x = element_text(color = font_color),
#            axis.text.y = element_text(color = font_color),
#            legend.title = element_text(color = font_color),
#            legend.text = element_text(color = font_color),
#            panel.grid.minor = element_line(color = grid_color_minor),
#            panel.grid.major = element_line(color = grid_color_major),
#            panel.background = element_rect(fill = back_color, color = font_color)
#          )
# 
# # I also want to see a residuals plot in multiples of SD of productivity to help spot outliers
# breaks_resid_sd_fig = ggplot( breaks_model_lin, aes(.fitted , .resid / breaks_descr$descriptives$asDF$`product[sd]` ) )
# 
# breaks_resid_sd_fig + 
#     geom_point(col = font_color) +
#     geom_hline(yintercept=0, col="green3", linetype="dashed") +
#     xlab("Fitted Productivity Values") +
#     ylab(expression( "Residuals - Standardized (Multiples of " * sigma * ")") )  +
#     ggtitle("Productivity Residual vs. Fitted Plot") +
#     theme_minimal() +
#     theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
#            axis.title.x = element_text(color = font_color),
#            axis.title.y = element_text(color = font_color),
#            axis.text.x = element_text(color = font_color),
#            axis.text.y = element_text(color = font_color),
#            legend.title = element_text(color = font_color),
#            legend.text = element_text(color = font_color),
#            panel.grid.minor = element_line(color = grid_color_minor),
#            panel.grid.major = element_line(color = grid_color_major),
#            panel.background = element_rect(fill = back_color, color = font_color)
#          )  
# 
# # Run a Breusch Pagan Test of homoscedasticity
# car::ncvTest(breaks_model_lin)
```

### Data Cleaning - OPTIONAL
If there are data issues that need to be adjusted, clean the data here.

``` {r data_cleaning_raw}

```

### Build Simple Regression Model

Centered data is already saved from scatter plots above (shift the y-intercept to the mean of the dataset).  
Now we can build the regression model.  

```{r simple_regression}

# SAMPLE CODE
# #simple regression looking at self-efficacy (centered) as a predictor of desire to lead in a group (e.g. self-efficacy (centered) regressed on desire to lead)
# breaks_model_lin_reg <- linReg( data = breaks_dat, 
#                                 dep = 'product', 
#                                 covs = c('length_centered'), 
#                                 blocks = list( c( 'length_centered') ), 
#                                 modelTest = TRUE, 
#                                 stdEst = TRUE,
#                                 ci = TRUE)
# breaks_model_lin_reg


```


