---
title: "PSYCH308B - Data Analysis 03 (DA03)"
author: "Brady C. Jackson"
date: "2024/11/10"

# Write document output to HTML, Word, and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

```{r package_loading, message=FALSE, warning=FALSE}

# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(jmv)       # for descriptive
library(ggplot2)
library(dplyr)
library(tidyr)       # for pivot_longer
# library(car)       # for leveneTest
# library(stringr)   # for sub_str operations
library(psych)
# library(Hmisc)     # for fun.dat substitution
# library(see)       # for outliers analysis 
# library(magrittr)
# library(AER)
library(cowplot)   # for axis aligned grid plots
library(afex)      # for pairwise confidence intervals
library(emmeans)   # for pairwise confidence intervals
library(pastecs)
```

## Investigation of Raw, Unprocessed Data

### Data Loading

First we need to load the data. 

```{r data_prep}

# Load in the data
dats_bodies_full <- read.csv("./Bernard et al. (2012).csv")

# Our research introduction doesn't tell us what "1" and "2" mean in the "gender" column so we can't
# add any data labels for that category. Notably, our research question doesn't rely on understanding the gender of the
# participants so we'll drop that column before doing further data sorting and preping
colnames(dats_bodies_full) <- tolower(colnames(dats_bodies_full))
dats_bodies <- subset(dats_bodies_full, select = -gender)

# We'll be sure to include some display name versions of the
# column headers in the long form data. So first let's create the longform data.
dats_bodies_long <- dats_bodies %>% tidyr::pivot_longer(!id, names_to = "pic_type", values_to = "recognition")

# Add a column of "display names" to the long-form data so we can have some print friendly output when needed
dats_bodies_long$pic_disp_names = dats_bodies_long$pic_type
dats_bodies_long$pic_disp_names[dats_bodies_long$pic_disp_names == "inverted_women"] <- 'Inverted Women'
dats_bodies_long$pic_disp_names[dats_bodies_long$pic_disp_names == "upright_women"] <- 'Upright Women'
dats_bodies_long$pic_disp_names[dats_bodies_long$pic_disp_names == "inverted_men"] <- 'Inverted Men'
dats_bodies_long$pic_disp_names[dats_bodies_long$pic_disp_names == "upright_men"] <- 'Upright Men'

# Cast both the display names and the pic_type columns as factors. Note that before-and-after arrays
# were manually inspected to ensure appropriate order of values were maintained for the cast arrays given
# Specifically, entries for ID==81 were inspected as that row has unique values for all four conditions
# of picture type.
dats_bodies_long$pic_type       <- factor( dats_bodies_long$pic_type, levels = c( "inverted_women",
                                                                                  "upright_women", 
                                                                                  "inverted_men", 
                                                                                  "upright_men" 
                                                                                 )
                                          )
dats_bodies_long$pic_disp_names <- factor(dats_bodies_long$pic_disp_names, levels = c( "Inverted Women",
                                                                                       "Upright Women", 
                                                                                       "Inverted Men", 
                                                                                       "Upright Men" 
                                                                                      )
                                          )
```

### Create Difference Scores

We need to add the difference scores to the base dataset (which already has gender removed as an unnecessary variable)


```{r calc_diff_scores}
# First we'll get the names of the independent variable columns from the longform data
pic_types = unique(as.character( dats_bodies_long$pic_type) )

# Now we loop over the list of names and compute difference scores to each name that is not the current one being
# differenced against. Once we difference each other column to the current pic_type, we pop that pic_type off so 
# that we don't compute differences against it again.
n_types = length(pic_types)
for(iii in 1:(n_types - 1) ){
    for(jjj in (iii + 1):n_types){
        this_type <- pic_types[iii]
        next_type <- pic_types[jjj]
        
        # Create the difference score column name
        diff_name <- paste(this_type, 'm', next_type, sep = '_')
        
        # Add the new difference score column to the non-long-wise data
        dats_bodies[[diff_name]] <- dats_bodies[[this_type]] - dats_bodies[[next_type]]
        
    }
}

```


### Formalize Model

We've been using the lm function to formalize a linear model thus far, but for this analysis we'll use the 
AOV function instead per Dr. Diaz's recommendations.

```{r raw_model_definition}

# We define our model as object recognition percentage (recognition) being dependent on pic_type.
bodies_model <- aov( recognition~pic_type, data=dats_bodies_long )

# We're also going to define a Long form dataframe for the difference scores (not the raw scores), and we'll define
# a second linear model that quantifies recognition_diff as a function of difference pairing
dats_bodies_diff <- subset(dats_bodies, select = -c(inverted_women, upright_women, inverted_men, upright_men))
dats_bodies_diff_long <- dats_bodies_diff %>% tidyr::pivot_longer(!id, names_to = "diff_pair", values_to = "recog_diff")

# We'll cast the longform diffs categories as factors and add a column of print-friendly factor names
# Add a column of "display names" to the long-form data so we can have some print friendly output when needed
dats_bodies_diff_long$diff_disp_names = dats_bodies_diff_long$diff_pair
dats_bodies_diff_long$diff_disp_names[dats_bodies_diff_long$diff_disp_names == "inverted_women_m_upright_women"] <- 'IW-UW'
dats_bodies_diff_long$diff_disp_names[dats_bodies_diff_long$diff_disp_names == "inverted_women_m_inverted_men"] <- 'IW-IM'
dats_bodies_diff_long$diff_disp_names[dats_bodies_diff_long$diff_disp_names == "inverted_women_m_upright_men"] <- 'IW-UM'
dats_bodies_diff_long$diff_disp_names[dats_bodies_diff_long$diff_disp_names == "upright_women_m_inverted_men"] <- 'UW-IM'
dats_bodies_diff_long$diff_disp_names[dats_bodies_diff_long$diff_disp_names == "upright_women_m_upright_men"] <- 'UW-UM'
dats_bodies_diff_long$diff_disp_names[dats_bodies_diff_long$diff_disp_names == "inverted_men_m_upright_men"] <- 'IM-UM'

# Cast both the display names and the pic_type columns as factors...I hate this, I need to get better about order
# control in factor vars.
dats_bodies_diff_long$diff_pair <- factor( dats_bodies_diff_long$diff_pair, levels = c( "inverted_women_m_upright_women",
                                                                                        "inverted_women_m_inverted_men", 
                                                                                        "inverted_women_m_upright_men",  
                                                                                        "upright_women_m_inverted_men",  
                                                                                        "upright_women_m_upright_men",   
                                                                                        "inverted_men_m_upright_men" 
                                                                                       )
                                          )
dats_bodies_diff_long$diff_disp_names <- factor(dats_bodies_diff_long$diff_disp_names, levels = c( "IW-UW",
                                                                                                   "IW-IM",
                                                                                                   "IW-UM",                                                                                                   
                                                                                                   "UW-IM",
                                                                                                   "UW-UM",
                                                                                                   "IM-UM"
                                                                                                  )
                                                )


# No we can define a second model as recognition differences dependent on difference pairs
bodies_diff_model <- aov( recog_diff~diff_disp_names, data=dats_bodies_diff_long )
```

### Data Descriptives Stats and Visualization



```{r raw_descriptives_and_visuals}

# Define font color for plots
font_color = "#4F81BD"

# Get JMV descriptives of raw scores
bodies_descr = jmv::descriptives( dats_bodies_long, 
                                  vars = "recognition", 
                                  splitBy="pic_disp_names", 
                                  hist = TRUE,
                                  dens = TRUE,
                                  qq = TRUE,
                                  sd = TRUE, 
                                  variance = TRUE,
                                  se = TRUE, 
                                  skew = TRUE, 
                                  kurt = TRUE,
                                 missing = TRUE)
print(bodies_descr)


# Get JMV descriptives of difference scores
bodies_diff_descr = jmv::descriptives( dats_bodies_diff_long, 
                                       vars = "recog_diff", 
                                       splitBy="diff_disp_names", 
                                       hist = TRUE,
                                       dens = TRUE,
                                       qq = TRUE,
                                       sd = TRUE, 
                                       variance = TRUE,
                                       se = TRUE, 
                                       skew = TRUE, 
                                       kurt = TRUE,
                                       missing = TRUE)
print(bodies_diff_descr)


# Try psych::describe. This should include descriptives for both raw scores and diff scores since we have both in the
# meshgrid version of the table
bodies_psych_descr <- psych::describe(dats_bodies, skew = TRUE)
print(bodies_psych_descr)


# Bar chart: We'll check the mean object recognition by picture type with 
#            confidence intervals included.
ggplot(dats_bodies_long, aes( pic_disp_names, recognition) ) +
    stat_summary(fun = mean, geom = "bar", position = "dodge", fill="steelblue3") +
    stat_summary( fun.data = mean_cl_normal, geom = "errorbar",
                  position = position_dodge(width = 0.90), width = 0.5
                 ) +
    labs( y = "Mean Object Recognition Fraciton",
          x = "Picture Type") +
    ggtitle('Impact of Picture Type on Object Recognition') +
    theme_minimal() +
    theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
           axis.title.x = element_text(color = font_color),
           axis.title.y = element_text(color = font_color)
          )
# Bar chart: We'll check the mean differences in object recognition by picture pairwise types with 
#            confidence intervals included.
ggplot(dats_bodies_diff_long, aes( diff_disp_names, recog_diff) ) +
    stat_summary(fun = mean, geom = "bar", position = "dodge", fill="forestgreen") +
    stat_summary( fun.data = mean_cl_normal, geom = "errorbar",
                  position = position_dodge(width = 0.90), width = 0.5
                 ) +
    labs( y = "Mean of Differences in Object Recognition Fraciton",
          x = "Picture Type Pairs") +
    ggtitle('Mean of Differences in Object Recognition Fraction') +
    theme_minimal() +
    theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
           axis.title.x = element_text(color = font_color),
           axis.title.y = element_text(color = font_color)
          )


# Category-wise Scatter Plot 
#  This shows us that the predictor variable is, indeed, categorical.
cat_plot = ggplot(dats_bodies_long, aes(x = pic_disp_names, y = recognition) ) + 
                  geom_point( aes(color = "Recognition Fraction") ) + 
                  labs(x = "Picture Type", y = "Recognition Fraction (0 - 1)") + 
                  ggtitle("Recognition Fraction binned by Picture Type") + 
                  geom_hline( aes( yintercept = 0, color = "Upper Limit" ), linetype = "dashed" ) + 
                  geom_hline( aes( yintercept = 1, color = "Lower Limit" ), linetype = "dashed" ) + 
                  scale_color_manual(
                                     name = "Legend",
                                     values = c( "steelblue3", "red4", "red4"),
                                     labels = c("Recognition Fraction", "Upper Limit", "Lower Limit")
                                    ) +
                 theme(legend.position = "right") +
    theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
           axis.title.x = element_text(color = font_color),
           axis.title.y = element_text(color = font_color)
          )
cat_plot



# Consider adding continuity plot:
# For one-way ANOVA's we explicitly checked continuity of data (e.g. output data generally spanned the entire
# range of allowed values). That wasn't listed as an assumption for RM ANOVA's by Dr. Diaz so I'm not plotting it here
# to save myself time. But you may consider adding later if you have time.



```



### Advanced Raw Visuals

More advanced visuals which can be based on model object go here.

```{r advanced_raw_visuals}

# Residuals histogram of raw data
resids_dats_bodies = bodies_model$residuals
hist( resids_dats_bodies, 
      main = "Histogram of Recognition Fraction Residuals", 
      xlab = "Residuals of Recognition Fractions",
      ylab = "Frequency"
     )


# Residuals histogram of difference scores
resids_dats_diffs = bodies_diff_model$residuals
hist( resids_dats_diffs, 
      main = "Histogram of Recognition Difference Scores Residuals", 
      xlab = "Residuals of Recognition Difference Scores",
      ylab = "Frequency"
     )
dats_bodies_diff_long$resids_dats_diffs <- resids_dats_diffs


# Groupwise residuals of difference scores using your DA02 method. I didn't do this for raw scores as it feels like
# overkill

# Now that we have a formalized model we can use the residuals command to output the residuals with respect to the
# mean separated by predictor-variable condition. We'll downselect the full
# array of residuals to subsets of residuals separated by haste condition
# NOTE: When applied to a formalized model listing a categorical variable as the predictor variable
#       (in this case, diff_disp_names), the residuals function computes the residuals with respect to the means of each
#       condition (level). So the residuals for difference scores are computed against the mean for a given difference score
resids_diffs_iw_uw  <- dats_bodies_diff_long$resids_dats_diffs[dats_bodies_diff_long$diff_disp_names == 'IW-UW']
resids_diffs_iw_im  <- dats_bodies_diff_long$resids_dats_diffs[dats_bodies_diff_long$diff_disp_names == 'IW-IM']
resids_diffs_iw_um  <- dats_bodies_diff_long$resids_dats_diffs[dats_bodies_diff_long$diff_disp_names == 'IW-UM']
resids_diffs_uw_im  <- dats_bodies_diff_long$resids_dats_diffs[dats_bodies_diff_long$diff_disp_names == 'UW-IM']
resids_diffs_uw_um  <- dats_bodies_diff_long$resids_dats_diffs[dats_bodies_diff_long$diff_disp_names == 'UW-UM']
resids_diffs_im_um  <- dats_bodies_diff_long$resids_dats_diffs[dats_bodies_diff_long$diff_disp_names == 'IM-UM']

# We need to pack all of the residuals subsets into dataframes to be usable by ggplot
df_iw_uw <- data.frame( diff_resids = resids_diffs_iw_uw )
df_iw_im <- data.frame( diff_resids = resids_diffs_iw_im )
df_iw_um <- data.frame( diff_resids = resids_diffs_iw_um )
df_uw_im <- data.frame( diff_resids = resids_diffs_uw_im )
df_uw_um <- data.frame( diff_resids = resids_diffs_uw_um )
df_im_um <- data.frame( diff_resids = resids_diffs_im_um )

# # We'll control x-limits so everything aligns nicely
# my_xlims = c(round( min(resids_dats_diffs) ) - 1,
#              round( max(resids_dats_diffsl) ) + 1)
# my_ylims = c(0, 1)
my_xlims = c(-0.75, 0.75)
my_ylims = c(0, 0.6)
bw = 0.1
# 
# IW-UW Residuals Histogram
iw_uw_resids_fig <- ggplot( df_iw_uw, aes(x = diff_resids) ) +
                            geom_histogram( aes(y = ..count.. / sum(..count..) ),  binwidth = bw, color = "steelblue3", fill = "steelblue3") +
                            geom_density( aes(y = ..density.. * bw), color = "steelblue4", size = 1) +
                            labs(x = "Recognition Fraction Differences", y = "Density") +
                            xlim(my_xlims) +
                            ylim(my_ylims) +
                            theme( axis.title.x = element_blank(),
                                   axis.title.y = element_text(color = font_color)
                                  ) + 
                            annotate("text", x = 0.4, y = 0.4, label = "IW - UW", color = "steelblue4", size = 3)


# print(iw_uw_resids_fig)

# IW-IM Residuals Histogram
iw_im_resids_fig <- ggplot( df_iw_im, aes(x = diff_resids) ) +
                            geom_histogram( aes(y = ..count.. / sum(..count..) ),  binwidth = bw, color = "purple3", fill = "purple3") +
                            geom_density( aes(y = ..density.. * bw), color = "purple4", size = 1) +
                            labs(x = "Recognition Fraction Differences", y = "Density") +
                            xlim(my_xlims) +
                            ylim(my_ylims) +
                            theme( axis.title.x = element_blank(),
                                   axis.title.y = element_text(color = font_color)
                                  ) + 
                            annotate("text", x = 0.4, y = 0.4, label = "IW - IM", color = "purple4", size = 3)

# print(iw_im_resids_fig)

# IW-UM Residuals Histogram
iw_um_resids_fig <- ggplot( df_iw_um, aes(x = diff_resids) ) +
                            geom_histogram( aes(y = ..count.. / sum(..count..) ),  binwidth = bw, color = "orange3", fill = "orange3") +
                            geom_density( aes(y = ..density.. * bw), color = "orange4", size = 1) +
                            labs(x = "Recognition Fraction Differences", y = "Density") +
                            xlim(my_xlims) +
                            ylim(my_ylims) +
                            theme( axis.title.x = element_blank(),
                                   axis.title.y = element_text(color = font_color)
                                  ) + 
                            annotate("text", x = 0.4, y = 0.4, label = "IW - UM", color = "orange4", size = 3)

# print(iw_um_resids_fig)

# UW-IM Residuals Histogram
uw_im_resids_fig <- ggplot( df_uw_im, aes(x = diff_resids) ) +
                            geom_histogram( aes(y = ..count.. / sum(..count..) ),  binwidth = bw, color = "orchid3", fill = "orchid3") +
                            geom_density( aes(y = ..density.. * bw), color = "orchid4", size = 1) +
                            labs(x = "Recognition Fraction Differences", y = "Density") +
                            xlim(my_xlims) +
                            ylim(my_ylims) +
                            theme( axis.title.x = element_blank(),
                                   axis.title.y = element_text(color = font_color)
                                  ) + 
                            annotate("text", x = 0.4, y = 0.4, label = "UW - IM", color = "orchid4", size = 3)

# print(uw_im_resids_fig)

# UW-UM Residuals Histogram
uw_um_resids_fig <- ggplot( df_uw_um, aes(x = diff_resids) ) +
                            geom_histogram( aes(y = ..count.. / sum(..count..) ),  binwidth = bw, color = "yellow3", fill = "yellow3") +
                            geom_density( aes(y = ..density.. * bw), color = "yellow4", size = 1) +
                            labs(x = "Recognition Fraction Differences", y = "Density") +
                            xlim(my_xlims) +
                            ylim(my_ylims) +
                            theme( axis.title.x = element_text(color = font_color),
                                   axis.title.y = element_text(color = font_color)
                                  ) + 
                            annotate("text", x = 0.4, y = 0.4, label = "UW - UM", color = "yellow4", size = 3)
# print(uw_um_resids_fig)

# IM-UM Residuals Histogram
im_um_resids_fig <- ggplot( df_im_um, aes(x = diff_resids) ) +
                            geom_histogram( aes(y = ..count.. / sum(..count..) ),  binwidth = bw, color = "firebrick3", fill = "firebrick3") +
                            geom_density( aes(y = ..density.. * bw), color = "firebrick4", size = 1) +
                            labs(x = "Recognition Fraction Differences", y = "Density") +
                            xlim(my_xlims) +
                            ylim(my_ylims) +
                            theme( axis.title.x = element_text(color = font_color),
                                   axis.title.y = element_text(color = font_color)
                                  ) + 
                            annotate("text", x = 0.4, y = 0.4, label = "IM - UM", color = "firebrick4", size = 3)
# print(im_um_resids_fig)


# Display residuals histograms in a subplot
resids_title = ggplot() +
               ggtitle("Recognition Fraction Pairwise Difference Score Residuals") +
               theme_void() +
               theme(plot.title = element_text(size = 12, face = "italic", color = font_color) )
resids_grid = plot_grid( iw_uw_resids_fig, iw_im_resids_fig, iw_um_resids_fig, uw_im_resids_fig, uw_um_resids_fig, im_um_resids_fig,
                         nrow = 3, ncol = 2,
                         align = "h"
                        )
resids_figure = plot_grid( resids_title, resids_grid,
                           ncol = 1,
                           rel_heights = c(0.05, 1)
                           )
print( resids_figure)


```

### Test / Analysis of Results with Mauchly's Test Included

Here we run the repeated measures analysis of variance. Mauchly's test, the final assumption check  
which is a check of sphericity across the data, is included as an ouput

```{r, rm_anova}

# To conduct the repeated measures ANOVA we can use the meshgrid version of the datatable but we need to use
# manually defined lists to clarify the levels of the predictor variable and which cells they are located in.
bodies_anova <- anovaRM(
  data = dats_bodies,
  rm = list( list( label = 'recog_diff', levels = c('inverted_women', 'upright_women', 'inverted_men', 'upright_men') ) ),
  rmCells = list( 
    list(measure = 'inverted_women', cell = 'inverted_women'),
    list(measure = 'upright_women', cell = 'upright_women'),
    list(measure = 'inverted_men', cell = 'inverted_men'),
    list(measure = 'upright_men', cell = 'upright_men')
  ),
  rmTerms = list('recog_diff'),
  effectSize = c('partEta', 'eta'),
  spherTests = TRUE,                          # Mauchly's test of sphericity: Recall - testing differences in variance of diff-scores
  spherCorr = c('none', 'GG'),                # Sphericity corrections: no correction and GG
  postHoc = list('recog_diff'),
  postHocCorr = 'holm',                       # Since repeated measure (don't use Tukey)
  emMeans = list('recog_diff'),
  emmTables = TRUE)

print(bodies_anova)


```

### Compute Confidence Intervales on Means of Difference Scores in Pairwise Comparisons

Since the anovaRM function doesn't output the confidence intervals on the pairwise comparisons, we do that
manually here.


```{r pairwise_effects_and_CI}

#
# Contrast function gives us a pairwise t-test (holm's)
# Summary w/ infer argument gives us Ci extracted from those holm's tests in Contrast
# Mutate gives us a manual calc of Cohen's d
# NOTE: comparisons are done in alphabetical order (base minus week3) so they're "negative" even though we should
#       report it as positive since time increases

#pairwise comparison results again but with confidence intervals and effect sizes, yay!
pairwise_bodies <- 
  emmeans(aov_ez(id = "id", dv = "recognition", within = "pic_type", data = dats_bodies_long), 
          pairwise ~ pic_type, adjust = "holm") %>%
  contrast(method = "pairwise", adjust = "holm") %>%
  summary(infer = TRUE) %>%
  mutate(Cohen_d = estimate / SE)

# Print results
print(pairwise_bodies)
```
