---
title: "PSY.308b.RMANOVA.DEMO"
output: html_document
---
 
```{r, echo=FALSE, results=FALSE, message=FALSE, cache=FALSE}
library(knitr); opts_chunk$set(error=TRUE, cache=FALSE)
```

```{r}
# Load packages
library(psych)
library(jmv)
library(ggplot2)
library(car)
library(pastecs)
library(dplyr)
library(magrittr)
library(tidyr)
library(afex)
library(emmeans)
```


EXAMPLE 1
A study conducted by Stiller-Ostrowski et al. (2005) assessed the effectiveness of an education intervention on injury psychology for athletic training students. Students (n = 11) completed a skill usage assessment at 5 timepoints (baseline, week 3, week 6, week 7, week 14) in order to determine the effectiveness of the intervention over time. 

IV: Training (Baseline, Week 3, Week 6, Week 7, Week 14)
DV: Skill Usage (1-300)

**SET UP**
```{r}
# Read data - notice it's in wide format
dat_injury <- read.csv("injury.csv")
```

```{r}
#changing to long for graphing purposes
dat_injury_long <- dat_injury %>% pivot_longer(!ID, names_to = "time", values_to = "usage")

#specifying the levels of time so they're in the right order when graphing
dat_injury_long$time <- factor(dat_injury_long$time,levels = c("base", "week3", "week6", "week7", "week14"))
```

```{r}
#check the class of all the variable to confirm
str(dat_injury_long)
```

**PREPARE: VISUALIZE (GET TO KNOW) YOUR DATA AND CHECK ASSUMPTIONS**
Assumption Reminder: The DV (residuals) of the DIFFERENCE scores must be normally distributed and the variances of the DIFFERENCE score distributions must be homogeneous (sphericity). We will check normality first looking at the histogram of the residuals and checking skew and kurtosis for all of the DIFFERENCE scores. For sphericity, we will run a Mauchly's test as a part of the RM ANOVA. 

```{R}
#creating a bar graph  
bar <- ggplot(dat_injury_long, aes(time, usage))

bar + stat_summary(fun = mean, geom = "bar", position = "dodge", fill = "tomato1") + stat_summary(fun.data = mean_cl_normal, geom = "errorbar", position = position_dodge(width = 0.90), width = 0.2) + labs(x = "Timepoint", y = "Skill Usage") + ggtitle('Effect of Time on Skill Usage')
```

```{r}
#Create difference scores 
dat_injury$base_week3 <- (dat_injury$base - dat_injury$week3)
dat_injury$base_week6 <- (dat_injury$base - dat_injury$week6)
dat_injury$base_week7 <- (dat_injury$base - dat_injury$week7)
dat_injury$base_week14 <- (dat_injury$base - dat_injury$week14)
dat_injury$week3_week6 <- (dat_injury$week3 - dat_injury$week6)
dat_injury$week3_week7 <- (dat_injury$week3 - dat_injury$week7)
dat_injury$week3_week14 <- (dat_injury$week3 - dat_injury$week14)
dat_injury$week6_week7 <- (dat_injury$week6 - dat_injury$week7)
dat_injury$week6_week14 <- (dat_injury$week6 - dat_injury$week14)
dat_injury$week7_week14 <- (dat_injury$week7 - dat_injury$week14)
```

```{r}
#checking normality graphically. We specify the model and then ask for a histogram of the model residuals (i.e. the amount the model over or under predicts each participant's skill usage score)
model.injury<-aov(usage~as.factor(time),data=dat_injury_long)
res.injury<-model.injury$residuals
hist(res.injury, main="Histogram of residuals",xlab="Residuals")
```

```{r}
#checking descriptive stats, including quantitative indicators of normality - specifically skew and kurtosis of the difference scores. 
desc.campaign.diff <- jmv::descriptives(dat_injury, vars = c('base_week3','base_week6','base_week7','base_week14','week3_week6', 'week3_week7','week3_week14', 'week6_week7', 'week6_week14','week7_week14'), sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.campaign.diff

#since we're checking so many variables, sometimes I like to use psych's describe function because it organizes the variables vertically rather than horizontally - can be helpful when there are a lot of variables
desc.campaign2 <- psych::describe(dat_injury)
desc.campaign2
```

**RM ANOVA**
```{r}
modelinjury <- anovaRM(
  data = dat_injury,
  rm = list(list(label = 'training', levels = c('base', 'week3', 'week6', 'week7', 'week14'))),
  rmCells = list(
    list(measure = 'base', cell = 'base'),
    list(measure = 'week3', cell = 'week3'),
    list(measure = 'week6', cell = 'week6'),
    list(measure = 'week7', cell = 'week7'),
    list(measure = 'week14', cell = 'week14')
  ),
  rmTerms = list('training'),
  effectSize = c('partEta', 'eta'),
  spherTests = TRUE,  # Mauchly's test of sphericity
  spherCorr = c('none', 'GG'),  # Sphericity corrections: no correction and GG
  postHoc = list('training'),
  postHocCorr = 'holm',
  emMeans = list('training'),
  emmTables = TRUE)

modelinjury

```

```{r}
#pairwise comparison results again but with confidence intervals and effect sizes, yay!
pairwise_injury <- 
  emmeans(aov_ez(id = "ID", dv = "usage", within = "time", data = dat_injury_long), 
          pairwise ~ time, adjust = "holm") %>%
  contrast(method = "pairwise", adjust = "holm") %>%
  summary(infer = TRUE) %>%
  mutate(Cohen_d = estimate / SE)

# Print results
print(pairwise_injury)
```


EXAMPLE 2
A local nonprofit is interested in increasing participation in their community volunteering program. They hire a market panel research company to test the effect of three different recruitment campaigns on the likelihood to join the program. The first campaign highlights the need, the second focuses on the impact volunteers have had in the past, and the third highlights the relationships and community that volunteers develop. They work with the market research firm to recruit a panel of 100 participants to review three pitches (one for each campaign) and rate their likelihood to volunteer after each campaign. 

IV: Campaign (Need, Impact, Relationship)
DV: Likelihood to Volunteer (1-25)

**SET UP**
```{r}
#Read data - notice the data is in long format
dat_vol <- read.csv("volunteer.csv")
```

```{r}
#check the class of all the variable
str(dat_vol)

#convert both subject and campaign to factors. We converted subject to factor so that R knows all of the 1s are one person and all the 2s are one person (but they aren't numerically related to one another). Subject is a categorical variable with 100 levels. Here we're using a different function. Mutate from dplyr is a common way to make all kinds of changes to variables. 
dat_vol <- dat_vol %>% dplyr::mutate_at(c('Subject', 'Campaign'), as.factor)
```

```{r}
#convert your data from long to wide format so you have it ready for the anovaRM function
dat_vol_wide <- pivot_wider(dat_vol, names_from = Campaign, values_from = Volunteer)
```

**PREPARE: VISUALIZE (GET TO KNOW) YOUR DATA AND CHECK ASSUMPTIONS**
Assumption Reminder: The DV (residuals) of the DIFFERENCE scores must be normally distributed and the variances of the DIFFERENCE score distributions must be homogeneous (sphericity). We will check normality first looking at the histogram of the residuals and checking skew and kurtosis for all of the DIFFERENCE scores. For sphericity, we will run a Mauchly's test as a part of the RM ANOVA. 

```{R}
#creating a bar graph  
bar <- ggplot(dat_vol, aes(Campaign, Volunteer))

bar + stat_summary(fun = mean, geom = "bar", position = "dodge", fill = "tomato1") + stat_summary(fun.data = mean_cl_normal, geom = "errorbar", position = position_dodge(width = 0.90), width = 0.2) + labs(x = "Campaign", y = "Mean Volunteering Likelihood") + ggtitle('Effect of Campaign Type of Likelihood to Volunteer')
```

```{r}

#descriptive split by campaign type to review means, sds, etc
desc.campaign <- descriptives(dat_vol, vars = c('Volunteer'), splitBy = 'Campaign',sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.campaign
```

```{r}
#checking normality graphically. We specify the model and then ask for a histogram of the model residuals (i.e. the amount the model over or under predicts each participant's skill usage score)
model.vol<-aov(Volunteer~as.factor(Campaign),data=dat_vol)
res.vol<-model.vol$residuals
hist(res.vol, main="Histogram of residuals",xlab="Residuals")
```

```{r}
#create difference scores to be able to check means an skew/kurtosis of those difference scores
dat_vol_wide$im_nd <- dat_vol_wide$Impact - dat_vol_wide$Need
dat_vol_wide$im_rel <- dat_vol_wide$Impact - dat_vol_wide$Relationship
dat_vol_wide$nd_rel <- dat_vol_wide$Need - dat_vol_wide$Relationship

desc.campaign.diff <- descriptives(dat_vol_wide, vars = c('im_nd', 'im_rel', 'nd_rel'), sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.campaign.diff
```

**RM ANOVA**
```{r}
modelvol <- anovaRM(data = dat_vol_wide, 
                 rm = list(list(label = 'Campaign', 
                                levels = c('Impact', 'Need', 'Relationship'))), 
                 rmCells = list(list(measure = 'Impact', cell = 'Impact'),
                                list(measure = 'Need', cell = 'Need'),
                                list(measure = 'Relationship', cell = 'Relationship')),
                 rmTerms = list('Campaign'),
                 effectSize = c('partEta', 'eta'),
                 spherTests = TRUE,
                 spherCorr = c('none','GG'),
                 postHoc = list('Campaign'),
                 postHocCorr = 'holm',
                 emMeans = list('Campaign'),
                 emmTables = TRUE)

modelvol
```

```{r}
#pairwise comparison results again but with confidence intervals and effect sizes, yay!
pairwise_volunteer <- 
  emmeans(aov_ez(id = "Subject", dv = "Volunteer", within = "Campaign", data = dat_vol), 
          pairwise ~ Campaign, adjust = "holm") %>%
  contrast(method = "pairwise", adjust = "holm") %>%
  summary(infer = TRUE) %>%
  mutate(Cohen_d = estimate / SE)

# Print results
print(pairwise_volunteer)
```





