---
title: "PSYCH308C - Data Analysis __ (DA__)"
author: "Brady C. Jackson"
date: "2025/__/__"

# Write document output to HTML, Word, and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

```{r package_loading, message=FALSE, warning=FALSE}

# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(jmv)       # for descriptive
library(ggplot2)
library(dplyr)
# library(car)       # for leveneTest
# library(stringr)   # for sub_str operations
# library(psych)
# library(Hmisc)     # for fun.dat substitution
# library(see)       # for outliers analysis 
# library(magrittr)
# library(AER)
```

## Investigation of Raw, Unprocessed Data

### Data Loading

First we need to load the data. 

```{r data_prep}

```

### Raw Data Descriptive Stats and Visualization

```{r raw_descriptives_and_visuals}


```

### Formalize Model

Model formalization for unprocessed data goes here.

```{r raw_model_definition}

```

### Advanced Raw Visuals

More advanced visuals which can be based on model object go here.

```{r advanced_raw_visuals}


```

### Test of Variance for Raw Data

[OPTIONAL] The following section checks the homogeneity of variance in all of   
the raw data. This may or may not be necessary depending on what kinds of models  
you're building.

```{r levenes_raw}

```

### Test / Analysis of Raw Results

Run the test / analysis you intend to conduct on the raw data (no outlier)  
manipulation for the record.

```{r, raw_t_test}

```

### Check Outliers Directly

[OPTIONAL] Can use check_outliers function here as necessary


```{r outlier_analysis}


```

### Decision to Handle Outliers

Use this section to drop or modify outliers.

```{r outlier_handling}


```

## Investigation of Processed Data

Replicate analysis post-outlier handling.

### Processed Data Descriptive Stats and Visualization

```{r processed_descriptives_and_visuals}

```

### Formalize Model of Processed Data

Formalizing the processed model.

```{r processed_model_definition}


```

### Advanced Processed Visuals

Processed data and model visualizations

```{r advanced_processed_visuals}


```

### Test of Variance for Processed Data

[OPTIONAL] Hederoscedasticity check for processed data.

```{r levenes_processed}

```

### Check Test / Analysis Results on Processed Data

Run the Test / Analysis on processed data

```{r, processed_t_test}


```

### Plot Residuals of Processed Data

Residuals plots of processed data.

```{r outlier_analysis_processed}

```

