---
title: "PSYCH308B - Data Analysis 05 (DA05)"
author: "Brady C. Jackson"
date: "2024/11/27"

# Write document output to HTML, Word, and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

```{r package_loading, message=FALSE, warning=FALSE}

# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(jmv)       # for descriptive
library(ggplot2)
library(dplyr)
library(car)       # for leveneTest
# library(stringr)   # for sub_str operations
# library(psych)
# library(Hmisc)     # for fun.dat substitution
# library(see)       # for outliers analysis 
# library(magrittr)
# library(AER)
library(pastecs)
```

## Investigation of Raw, Unprocessed Data

### Data Loading

First we need to load the data. 

```{r data_prep}
# Load data
dats_driving = read.csv("./308B.Data.DA5.csv")


# We're going to create print-friendly name columns for both convo difficulty
# and driving difficulty. This will enable both better figures and more useful
# sorting in model definition later.
dats_driving$convo_names <- dats_driving$convo
dats_driving$drive_names <- dats_driving$drive

# Perform the string substitutions here, first for the convo difficulty names
# second for the driving difficulty names
dats_driving$convo_names[dats_driving$convo_names == 'C'] <- "Control"
dats_driving$convo_names[dats_driving$convo_names == 'E'] <- "Easy"
dats_driving$convo_names[dats_driving$convo_names == 'D'] <- "Difficult"

dats_driving$drive_names[dats_driving$drive_names == 'L'] <- "Low"
dats_driving$drive_names[dats_driving$drive_names == 'M'] <- "Medium"
dats_driving$drive_names[dats_driving$drive_names == 'D'] <- "Difficult"

# Convert all "group" categories to factors
dats_driving$convo <- factor(dats_driving$convo, levels = c("C", "E", "D") )
dats_driving$convo_names <- factor(dats_driving$convo_names, levels = c("Control", "Easy", "Difficult") )

dats_driving$drive <- factor(dats_driving$drive, levels = c("L", "M", "D") )
dats_driving$drive_names <- factor(dats_driving$drive_names, levels = c("Low", "Medium", "Difficult") )

```


### Formalize Model

We formalize the model using the aov (analysis of variance) model before we  
develop our descriptive stats as handling the model object will make the descriptives  
plots and stats easier to produce in this case.


```{r raw_model_definition}

# We use R's built-in aov function to define our model as driving errors being dependent  
# on the interaction (product of) conversation and driving difficulty levels (3x3)
driving_model <- aov( errors ~ convo_names * drive_names, data = dats_driving )
model.tables(driving_model, type = "means")

```


### Raw Data Descriptive Stats and Visualization

We'll make some basic visuals and descriptives tables here to help use visualize and analyze our data

```{r raw_descriptives_and_visuals}

# First we'll defines some meta-data to use in all of our plots so they're nice and clean
font_color = "#4F81BD"
grid_color_major = "#9BB7D9"
grid_color_minor = "#C8D7EA"
back_color = "gray95"

# Use "by" function to get descriptives
by(dats_driving$errors, list(dats_driving$drive_names, dats_driving$convo_names), basic = FALSE, norm = TRUE, stat.desc)


# Bar chart: We'll check the mean driving errors by driving condition with 
#            confidence intervals included.
drive_bar_hand <- ggplot(dats_driving, aes( drive_names, errors, fill = convo_names) ) +
                         stat_summary( fun = mean, geom = "bar", position = "dodge") +
                         stat_summary( fun.data = mean_cl_normal, geom = "errorbar",
                                       position = position_dodge(width = 0.90), width = 0.5
                                      ) +
                         scale_fill_manual( values = c("hotpink4", "seagreen3", "darkorange3") ) + 
                         labs( y = "Mean Driving Errors Committed",
                               x = "Driving Difficulty Level",
                               fill = "Conversation Difficulty") +
                         ggtitle('Driving Errors by Conversation & Driving Difficulty Condition') +
                         theme_minimal() +
                         theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
                                axis.title.x = element_text(color = font_color),
                                axis.title.y = element_text(color = font_color),
                                   axis.text.x = element_text(color = font_color),
                                   axis.text.y = element_text(color = font_color),
                                   legend.title = element_text(color = font_color),
                                   legend.text = element_text(color = font_color),
                                   panel.grid.minor = element_line(color = grid_color_minor),
                                   panel.grid.major = element_line(color = grid_color_major),
                                   panel.background = element_rect(fill = back_color, color = font_color)
                               )

print(drive_bar_hand)

# Line Chart: Same divisions / levels as bar chart, but in line form
drive_lines_hand <- ggplot( dats_driving, aes( drive_names, errors, color = convo_names) ) + 
                            stat_summary(fun = mean, geom = "point") + 
                            stat_summary(fun = mean, geom = "line", aes(group = convo_names) ) + 
                            stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) +
                            scale_color_manual("Conversation Difficulty", values = c("hotpink4", "seagreen3", "darkorange3") ) + 
                            labs( y = "Mean Driving Errors Committed",
                                  x = "Driving Difficulty Level",
                                  color = "Conversation Difficulty") +
                            ggtitle('Driving Error Trends by Conversation & Driving Difficulty Condition') +
                            theme_minimal() +
                            theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
                                   axis.title.x = element_text(color = font_color),
                                   axis.title.y = element_text(color = font_color),
                                   axis.text.x = element_text(color = font_color),
                                   axis.text.y = element_text(color = font_color),
                                   legend.title = element_text(color = font_color),
                                   legend.text = element_text(color = font_color),
                                   panel.grid.minor = element_line(color = grid_color_minor),
                                   panel.grid.major = element_line(color = grid_color_major),
                                   panel.background = element_rect(fill = back_color, color = font_color)
                                  )

print(drive_lines_hand)



```

### Assumption Checks

We'll do our formal assumption checks here

```{r advanced_raw_visuals}

# More Plot Metadata
my_xlims = c(-10, 10)
my_ylims = c(0, 0.2)
bw = 2


# Residuals Histogram
model_df <- data.frame(driving_resids = driving_model$residuals)
#hist(model_df$driving_resids, main = "Histogram of Residuals", xlab = "residuals")

resids_fig <- ggplot( model_df, aes(x = driving_resids) ) +
                      geom_histogram( aes(y = ..count.. / sum(..count..) ),  binwidth = bw, color = "purple3", fill = "purple3") +
                      geom_density( aes(y = ..density.. * bw), color = "purple4", size = 1) +
                      labs(x = "Driving Errors", y = "Density") +
                      ggtitle('Histogram of Driving Residuals Error') +
                      xlim(my_xlims) +
                      ylim(my_ylims) +
                      theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
                                   axis.title.x = element_text(color = font_color),
                                   axis.title.y = element_text(color = font_color),
                                   axis.text.x = element_text(color = font_color),
                                   axis.text.y = element_text(color = font_color),
                                   legend.title = element_text(color = font_color),
                                   legend.text = element_text(color = font_color),
                                   panel.grid.minor = element_line(color = grid_color_minor),
                                   panel.grid.major = element_line(color = grid_color_major),
                                   panel.background = element_rect(fill = back_color, color = font_color))
print(resids_fig)


# Test homgeneity of variance using Levene's test.
leveneTest(dats_driving$errors, interaction(dats_driving$drive_names, dats_driving$convo_names), center = mean)

```

### Test / Analysis of Raw Results

We can use JMV's anova function on our data to check main effects and interactive effects  
simultaneously.

```{r, raw_test}

jmv::ANOVA( data = dats_driving,
            dep = 'errors',                                   # ANOVA - Evaluate helping likelihood as outcome var
            factors = c('drive_names', 'convo_names'),        # ANOVA - split by haste condition
            effectSize = 'partEta',                           # ANOVA - report fraction of variance, not Cohen's d
            postHoc = c('drive_names', 'convo_names'),        # POST HOC - Also split post-hoc's by haste condition
            postHocCorr = c('tukey'),                         # Include both Tukey's and Bonferroni's
            postHocES = 'd',                                  # POST HOC
            emMeans = list('drive_names', 'convo_names'),     # DESCRIPTIVES BECAUSE JESS IS LAZY
            emmTables = TRUE,                                 # YES, MAKE ME A TABLE BECAUSE I'M LAZY
            emmPlotError = 'ci',
            emmPlotData = TRUE,
            homo = TRUE
           )

```

### Simple Effects Tests

We're going to need to run 3x ANOVAs for our simple effects tests. We'll subset by  
driving difficulty level so that we're evaluating the perspective as:  

  For a given driving difficult level, what's the impact of participating in a given  
  conversation difficulty level?  

```{r simp_effects}

# Subset the data by driving difficulty level
dats_drive_low <- subset(dats_driving, dats_driving$drive_names == "Low")
dats_drive_med <- subset(dats_driving, dats_driving$drive_names == "Medium")
dats_drive_dif <- subset(dats_driving, dats_driving$drive_names == "Difficult")

# ANOVA for Low Driving Difficulty Level
jmv::ANOVA( data = dats_drive_low,
            dep = 'errors',                                   # ANOVA - Evaluate helping likelihood as outcome var
            factors = c('convo_names'),                       # ANOVA - split by condition
            effectSize = 'partEta',                           # ANOVA - report fraction of variance, not Cohen's d
            postHoc = c('convo_names'),                       # POST HOC - Also split post-hoc's by haste condition
            postHocCorr = c('tukey'),                         # Include both Tukey's and Bonferroni's
            postHocES = 'd',                                  # POST HOC
            emMeans = c('convo_names'),                       # DESCRIPTIVES BECAUSE JESS IS LAZY
            emmTables = TRUE,                                 # YES, MAKE ME A TABLE BECAUSE I'M LAZY
            emmPlotError = 'ci',
            emmPlotData = TRUE,
            homo = TRUE
           )


# ANOVA for Medium Driving Difficulty Level
jmv::ANOVA( data = dats_drive_med,
            dep = 'errors',                                   # ANOVA - Evaluate helping likelihood as outcome var
            factors = c('convo_names'),                       # ANOVA - split by condition
            effectSize = 'partEta',                           # ANOVA - report fraction of variance, not Cohen's d
            postHoc = c('convo_names'),                       # POST HOC - Also split post-hoc's by haste condition
            postHocCorr = c('tukey'),                         # Include both Tukey's and Bonferroni's
            postHocES = 'd',                                  # POST HOC
            emMeans = c('convo_names'),                       # DESCRIPTIVES BECAUSE JESS IS LAZY
            emmTables = TRUE,                                 # YES, MAKE ME A TABLE BECAUSE I'M LAZY
            emmPlotError = 'ci',
            emmPlotData = TRUE,
            homo = TRUE
           )

# ANOVA for Difficult Driving Difficulty Level
jmv::ANOVA( data = dats_drive_dif,
            dep = 'errors',                                   # ANOVA - Evaluate helping likelihood as outcome var
            factors = c('convo_names'),                       # ANOVA - split by condition
            effectSize = 'partEta',                           # ANOVA - report fraction of variance, not Cohen's d
            postHoc = c('convo_names'),                       # POST HOC - Also split post-hoc's by haste condition
            postHocCorr = c('tukey'),                         # Include both Tukey's and Bonferroni's
            postHocES = 'd',                                  # POST HOC
            emMeans = c('convo_names'),                       # DESCRIPTIVES BECAUSE JESS IS LAZY
            emmTables = TRUE,                                 # YES, MAKE ME A TABLE BECAUSE I'M LAZY
            emmPlotError = 'ci',
            emmPlotData = TRUE,
            homo = TRUE
           )

```
