---
title: "Moderation Demo"
author: "Jessica Diaz"
date: "2/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(car)
library(psych)
library(jmv)
library(ggplot2)
library(tidyverse)
library(magrittr)
library(broom)
library(GGally)
library(foreign)
library(ltm)
library(mvnTest)
library(rockchalk)
library(haven)
library(jtools)
library(interactions)
# library(pequod)
```

**Continuous Moderation**

You are a positive social psychologist studying the effect of burnout on positive coping strategies. Your past research demonstrates that the most burned out an individual is, the less they are able to exhibit positive coping strategies in the face of challenge. However, you've noticed that this isn't a consistent reaction - some individuals seem to be more resistant to burnout's impact than others. You have a suspicion that personality, specifically extraversion may play a role in this difference. You decide to test this hypothesis

Variables
extrav: Extraversion
neur: Neuroticism
burn: Burnout
poscop: Use of positive coping strategies
negcop: Use of negative coping strategies
anxiety: state anxiety

RQ: Does extraversion moderate the relationship between burnout and use of positive coping  strategies?
```{r}
#Read in your data. As long as its in the same folder as the RMD file, this is the only code you need. 
dat.coping<- haven::read_dta("coping.dta")

#convert to lower case for ease of entering in variable names
colnames(dat.coping) %<>% tolower()
```

```{r}
#descriptives of your data
desc.coping <- descriptives(data = dat.coping, hist = TRUE, sd = TRUE, range = TRUE, skew = TRUE, kurt = TRUE)
desc.coping
```

```{r}
#listwise deleting participants with missing data
dat.coping.comp <- na.omit(dat.coping)
```

```{r}
#scatterplots for each predictor and your outcome variable
scatter.coping.burn <- ggplot(dat.coping.comp, aes(burn, poscop))
scatter.coping.burn + geom_point() + geom_smooth(method = "lm", colour = "Blue", se = FALSE) + labs(x = "Burnout", y = "Positive Coping")

scatter.coping.extrav <- ggplot(dat.coping.comp, aes(extrav, poscop))
scatter.coping.extrav + geom_point() + geom_smooth(method = "lm", colour = "Blue", se = FALSE) + labs(x = "Extraversion", y = "Positive Coping")
```

```{r}
#correlation table
cor.coping <- corrMatrix(dat.coping.comp)
cor.coping
```

```{r}
#create a saturated models of correlated variables to test normality and homoscedascity
model.coping.sat <- lm(poscop~burn+extrav, data=dat.coping.comp)

#HOMOSCEDASTICITY
#plot fitted values v. residuals to examine homoscedasticity
ggplot(model.coping.sat, aes(.fitted, .resid))+geom_point()+geom_hline(yintercept=0, col="red", linetype="dashed")+xlab("Fitted values")+ylab("Residuals")+ggtitle("Residual vs Fitted Plot")+theme_bw()

#test of homoscedasticity
ncvTest(model.coping.sat)
```

```{r}
#MULTIVARIATE NORMALITY
#qqplot to look at multivariate normality
ggplot(model.coping.sat, aes(qqnorm(.stdresid)[[1]], .stdresid))+geom_point(na.rm = TRUE)+xlab("Theoretical Quantiles")+ylab("Standardized Residuals")+ggtitle("Normal Q-Q")

#Henze-Zingler test of multivariate normality
HZ.test(dat.coping.comp[1:6])
```

```{r}
#Center our predictors for interpretability and to assist with multicollinearity when we make a product term
dat.coping.comp$burn.c <- dat.coping.comp$burn - mean(dat.coping.comp$burn, rm.na=TRUE)
dat.coping.comp$extrav.c <- dat.coping.comp$extrav - mean(dat.coping.comp$extrav, rm.na=TRUE)
```

```{r}
#Create a moderator term (product term)
dat.coping.comp$mod <- dat.coping.comp$burn.c*dat.coping.comp$extrav.c
```

```{r}
#Hierarchical Linear Regression
model.coping <- linReg(data = dat.coping.comp, 
                 dep = 'poscop', 
                 covs = c('burn.c', 'extrav.c', 'mod'), 
                 blocks = list(
                   list('burn.c','extrav.c'),
                   list('mod')), 
                modelTest = TRUE,
                collin = TRUE, 
                stdEst = TRUE,
                r2Adj = TRUE, 
                ci = TRUE)
model.coping
```

```{r}
#Unstandardized Simple Slopes Analysis

library(jtools)
library(interactions)
library(effectsize)

# Fit model with raw variables
model.poscomp <- lm(poscop ~ burn * extrav, data = dat.coping.comp)

# Now run simple slopes 
# NOTE: For these functions, we have to explicitly note who our perdictor is and who our moderator is
sim_slopes(model.poscomp, pred = "burn", modx = "extrav")

# Plot interaction
interact_plot(model.poscomp, pred = "burn", modx = "extrav")
```

```{r}
#Standardized Simple Slopes Analysis

library(jtools)
library(interactions)
library(effectsize)

# Fit model with raw variables
model.poscomp <- lm(poscop ~ burn * extrav, data = dat.coping.comp)

# Fully standardize model automatically
# NOTE: This generates the same thing as above, but for a centered (standardized model)
model.poscomp_std <- standardize(model.poscomp)

# Now run simple slopes 
sim_slopes(model.poscomp_std, pred = "burn", modx = "extrav")

# Plot interaction
interact_plot(model.poscomp_std, pred = "burn", modx = "extrav")
```

