---
title: "PSYCH308A - Data Analysis 4 (DA4)"
author: "Brady C. Jackson"
date: "2024/10/04"

# Write document output to both HTML and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  pdf_document:
    toc: true

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

```{r package_loading, message=FALSE, warning=FALSE}

# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(psych)
library(jmv)
library(ggplot2)
library(dplyr)
library(magrittr)
```

---

# Question 01 (Q01):

01. Interpret a p-value of .042. (This question is **not** asking for the  
    decision this p-value results in, rather what does this value mean?)

## Answer to Q01: 

A p-value of .042 implies that the finding of some statistical test (z-test,  
t-test, etc.) that the probability of obtaining the test statistic found or  
something more extreme was .042 (a.k.a. 4.2%) assuming the null hypothesis was  
true

---

# Question 02 (Q02):

2. ***In no more than two sentences,*** what is the relationship between sample  
   size, effect size, and power?

## Answer to Q02: 

Power is positively correlated with both sample size and effect size in a  
manner that convolutes the two (sample and effect). That is to say, the power  
of a finding will increase if either sample size increases or effect size  
increases, and if one (e.g. sample size) is necessarily small, a higher  
power can be achieved by increasing the other (e.g. effect size).

---

# Question 03 (Q03):

03. A researcher records the number of words recalled by students presented with  
    a list of words for 1 minute. In one group, students were presented with the  
    list of words in color; in a second group, the same words were presented in  
    black and white. An equal number of students were in each group. The  
    researcher reports the following: *Participants recalled significantly more  
    words when the words were presented in color (M = 12.4 words) versus black  
    and white (M = 10.9 words), t(48) = 2.01, p = .035, d = 0.18.* Based on the  
    previous statement, what is the sample size in each group?

## Answer to Q03: 
Given the wording of the prompt, I'm assuming this test was conducted as an  
independent sample t-test ("In one group ... in a second group" instead of  
phrasing like, "...the group of students looked at colored words and then  
later looked at black and white words...").  

The degrees of freedom for the t-test ran is given as 48 ( *t(48)* ).  
As this is an indepdendent t-test, two parameters had to be estimated, the  
mean of the population from which group 1 was drawn, and the mean of the  
population from which group 2 was drawn. Therefore, we have the relationship  
$$
\begin{aligned}
& df\ =\ [Things\ we\ know\ (n_{total}) ] - [Estimated\ Parameters\ (n_{P,est})] \\  
& df\ =\ n_{total}\ -\ n_{P,est} \\
& 48\ =\ n_{total}\ -\ 2 \\
& 48\ +\ 2\ =\ n_{total} \\
& 50 =\ n_{total} 
\end{aligned}
$$

Therefore, there were a total of 50 students in the study, with 25 in each group  
given that the prompt states the same number of students were in each group.

---

# Research Prompt for Q04 through Q09:

You are teaching your first Intro to Psychology course! After the midterm, you  
are disappointed with your students’ overall test scores. You decide to   
implement two different required study techniques. There are 100 students in   
the class; 50 of them will be required to meet in groups to study right before   
the final (Group A) and the other 50 will be required to create flashcards to   
aid in memorization (Group B). You are interested in two primary research   
questions:

---

## Research Question 1 (RQ1): Q04 - Q06:

**Did student test scores improve significantly from the midterm to the final?**  
**Data:** *308A.RQ1 Data.DA4.csv*

---

### Question 04 (Q04):

04. Visualize your data for this research question. Include your visualization  
    here.

#### Answer to Q04:

##### Loading Data for Q04

The first step to visualizing the data is to load it. See code below:  

```{r, q04_data_load}

# In order to visualize the data we must first load it. To do so we lool
# in the present working directory for CSV files, and take the one that
# has RQ1 in its filename. Append it to the current working directory
# to create the fullpath for loading
here <- getwd()
rq1_name <- list.files(here, pattern = ".*RQ1.*csv")

# Use the file.path function to create a platform appropriate fullpath / filename
# to the Research Question 1 data.
rq1_file <- file.path(here, rq1_name)

# Read the CSV data in as rq1_dat
rq1_dat <- read.csv(rq1_file, header = TRUE)

# Lower-case all column names for convenience
colnames(rq1_dat) <- tolower(colnames(rq1_dat))

```

##### Data Prep Q04

Ultimately we're going to answer Research Question 1 as a dependent T-test   
since we're evaluating change in student test scores over time (regardless)   
of study method. So we're going to prepare the data further with some math.  
See code below:  

```{r, q04_data_prep}
# Since we want to know if scores improved between the midterm and final we'll
# look at midterm scores, final scores, and the delta between the two. 
# To do this, we need to add one more column to our math, the difference
# between the two. We need to maintain chirality (direction) as it will
# be informative if scores rise or decrease so we will do final - midterm
# without taking the absolute value. If scores decreased, we should see a 
# positive mean of the diff. If they decreased we'll see a negative mean
# of the diff.
rq1_dat$fin_mid_diff <- rq1_dat$final - rq1_dat$midterm
```

##### Visualize Prepared Data Q04 - Descriptives

See the histograms below for visualization of the data pertinent to this  
research question. Bar graphs of the midterm scores, the final scores, and 
the diff of final - midterm are also shown for better understanding

```{r, q04_answer_and_visualization_outputs}

# We're going to create a descriptives object so we can checkout the histogram
# and descriptive stats of the midterm scores, the final scores, and the
# diff of both.
rq1_desc <- jmv::descriptives( rq1_dat[2:4], 
                               hist = TRUE,
                               dens = TRUE,
                               sd = TRUE,
                               variance = TRUE,
                               se = TRUE, 
                               skew = TRUE, 
                               kurt = TRUE
                             )

# Render the rq1 descriptives object as output.
rq1_desc

# Now we're going to create a bar chart of the midterm scores, final scores,
# and diffs .. though the diffs data willbe oddly scaled (much smaller) than the
# raw scores

# We're going to lay all three datasets out on a single bar graph 
# even though that may jackup the scaling to be bad for the diff.
bar_rq1 <- ggplot(rq1_dat, aes(colnames(rq1_dat)[2:4], rq1_dat[2:4]))



#make a bar graph of relationship between fieldtrip and hard tangrams
# START HERE FIX BUGGY CODE BELOW
# bar_rq1 + stat_summary( fun = mean, 
#                         geom = "bar", 
#                         position = "dodge", 
#                         fill="slateblue1"
#                       ) + 
#           stat_summary( fun.data = mean_cl_normal, 
#                         geom = "errorbar", 
#                         position = position_dodge(width = 0.90), 
#                         width = 0.2
#                       ) + 
#           labs(x = "Test Type", y = "Score (pts))") + 
#           ggtitle('Test Scores and Difference')

```



---

### Question 05 (Q05):

05. Did student test scores improve significantly from the midterm to the final?  
    Using RStudio to analyze, conduct a hypothesis test to evaluate this  
    question. Organize your answer according to the 4 steps of hypothesis  
    testing.

#### Answer to Q05: 
This is my answer

---

### Question 06 (Q06):

06. Report your findings in APA format. (Hint: make sure to answer the research  
    question!)

#### Answer to Q06: 
This is my answer

---

## Research Question 2 (RQ2): Q07 - Q09:

**Does the study technique used predict scores on the final exam? **  
**Data:** *308A.RQ2 Data.DA4.csv*

---

### Question 07 (Q07):

07. Visualize your data for this research question. Include your visualization  
    here.

#### Answer to Q07: 
This is my answer

---

### Question 08 (Q08):

08. Does the study technique used predict scores on the final exam? Using  
    RStudio to analyze, conduct a hypothesis test to evaluate this question.  
    Organize your answer according to the 4 steps of hypothesis testing.

#### Answer to Q08: 
This is my answer

---

### Question 09 (Q09):

09. ...

    a. Report your findings in APA format. (Hint: make sure to answer the  
       research question!)
    
    b. The Dean of the university was also interested in your results, as this  
       may help to raise scores in other departments. Unfortunately, she does  
       not understand statistical language. Please interpret your findings for  
       the Dean. Did scores improve? Which technique is better?

#### Answer to Q09: 
This is my answer

---

# Question 10 (Q10):

10. A developmental psychologist is interested in the effect of a positive  
    psychology intervention on the well-being of aging adults. She administers  
    the intervention, collects well-being scores from a sample of 100  
    participants, and tests whether their well-being differs significantly from  
    the national average. Using G*Power, she determines the power for her test  
    is .80. 

    a. Interpret this value 
    
    b. What suggestion would you give her if she wants a higher probability of  
       detecting a true effect?

## Answer to Q10: 
This is my answer

---

# Question 11 (Q11):

11. What would it mean if your analysis returned the following values? Consider   
    the meaning of t - not the decision associated with it.

    a. *t(24) = 0.35*
    
    b. *t(24) = 1.00*
    
    c. *t(24) = 3.2*

## Answer to Q11: 
This is my answer

---

# Question 12 (Q12):

12. Draw and annotate all the properties of the null and alternative curves:   
    power, beta, alpha, type 1 error, type 2 error. 

## Answer to Q12: 
This is my answer


