---
title: "PSYCH308C - Data Analysis 01 (DA01)"
author: "Brady C. Jackson"
date: "2025/01/23"

# Write document output to HTML, Word, and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prompt
A local business has implemented a new program to encourage employees to take more control over their workday. According   
to the program, employees are allowed to take their break at any point during the day for as long as they want between 1   
and 60 minutes. As businesses do, they were wondering how this program may relate to productivity on a 1-100 scale. At   
the end of the first month, the business conducted a short survey which had all 205 employees report the average length   
of their break. The business employed you to investigate how break length may relate to productivity.

## Variables 

| Variable | Type | Description |
|:-----|:---|:------------------|
| `Length` | numeric | length of break in minutes (range 1-60). |
| `Enjoy` | numeric | self-reported level of enjoyment of break period (scale average; range 1-10). |
| `Product (Productivity)` | numeric | percentage of time meeting weekly goals (range 1-100; represented as whole number). |

## Assignment

### Part 1

Run the appropriate assumptions and analyses and report your findings in an APA format results section.   
Include tables and figures as necessary. 

### Part 2

*Respond to the following questions.*

1. Share your findings with the owner of the business. Be sure to provide a recommendation informed by your results.  
2. Condense your findings and share them as you would to a statistics-saavy audience.   
3. If someone took an average-length break, what is their predicted productivity score?  
4. Find one empirical article in your concentration that uses some form of regression. Please read through and outline 
   the information reported in the results section only. Please include an APA format citation of the article you chose.

Please submit both your response to the above prompts AND your knitted code on Canvas. 

---ONLY WRITE CODE BELOW THIS LINE---

## Code

### Libraries

Load all requisite libraries here.

```{r package_loading, message=FALSE, warning=FALSE}
# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(jmv)       # for descriptive
library(ggplot2)
library(dplyr)
library(corrplot)    # For fancy covariance matrix plots
library(apaTables)   # For Word formatted tables
library(car)         # for ncvTest (Breusch Pagan)
# library(stringr)   # for sub_str operations
# library(psych)
# library(Hmisc)     # for fun.dat substitution
# library(see)       # for outliers analysis 
# library(magrittr)
# library(AER)
```

### Metadata

This section of code is to setup some general variables that we'll use throughout the code (e.g. figure colors, etc)

```{r metadata}
# First we'll defines some meta-data to use in all of our plots so they're nice and clean
font_color = "#4F81BD"
grid_color_major = "#9BB7D9"
grid_color_minor = "#C8D7EA"
back_color = "gray95"
rb_colmap = colorRampPalette( c("firebrick", "grey86", "dodgerblue3") )(200)


```

### Load and View data

```{r load_raw_data}
# Stuff goes here
breaks_raw_dat = read.csv("./da01.csv")

# Rename columns to lower because why not
colnames(breaks_raw_dat) <- tolower( colnames(breaks_raw_dat) )

# View data
glimpse(breaks_raw_dat)

# Ensure that the employee numbers of each subject in the study are unique to prevent any duplicate data
# if the size of the unique-entries only is the same as the whole vector then there are no duplicate subjects
test_unique = (length(unique(breaks_raw_dat$subject)) == length(breaks_raw_dat$subject))
if(!test_unique){
    print("WARNING: There are duplicate data entries in the raw data")
}else{
    print("No duplicate entries detected in raw data")
}
```

### Descriptive Statistics - Raw Data
This section will look at base descriptive statistics of the raw data to help identify data anomalies and check   
normality of predictor variables (break length - minutes)

```{r descriptive_stats_raw}
# Get JMV descriptives of raw productivity, enjoyment, and break length
# Tabulate mean, variance, skew, and kurtosis so we can check normality.
# Output histograms and qq plots so we can spot outliers and review distribution shape
breaks_descr = jmv::descriptives( breaks_raw_dat[c(2:4)],
                                  hist = TRUE,
                                  dens = TRUE,
                                  qq = TRUE,
                                  sd = TRUE, 
                                  variance = TRUE,
                                  se = TRUE, 
                                  skew = TRUE, 
                                  kurt = TRUE,
                                  missing = TRUE)
print(breaks_descr)

```

### Correlation Plots - Raw Data
Visualize the covariance matrix to understand correlation between break length and employee productivity


```{r correlation_plots_raw}
# Given that we're working explicitly on break length and productivity, I want to create a second dataframe that 
# consists only of those columns for simplicity. I'll maintain the original as well so I can get some insight into
# employee enjoyment as well, but since that's not the focus I want a clean dataset as well.
breaks_dat <- breaks_raw_dat
breaks_dat$subject <- NULL
breaks_dat$enjoy <- NULL

# Produce a correlation matrix plot that includes the numbers printed with a color gradient for easy reading.
# par(bg = "lightgrey")
breaks_full_cor <- stats::cor(breaks_raw_dat)
corrplot( breaks_full_cor, 
          method="color", 
          type = "full", 
          addCoef.col = "black", 
          col = rb_colmap, 
          tl.col = font_color )

breaks_short_cor <- stats::cor(breaks_dat)
corrplot( breaks_short_cor, 
          method="color", 
          type = "full",  
          addCoef.col = "black", 
          col = rb_colmap,  
          tl.col = font_color )


# Print correlation tables (annoying that we have to use two different correlation functions, 1 for plotting and
# 1 for tables, but oh well)
breaks_full_cor_tab <- jmv::corrMatrix(breaks_raw_dat[2:4], flag = TRUE)
print(breaks_full_cor_tab)

breaks_short_cor_tab <- jmv::corrMatrix(breaks_dat, flag = TRUE)
print(breaks_short_cor_tab)

# Finally dump a correlation table to Word Doc format for easy table formatting / to save time in writing
# the assignment
apa.cor.table( breaks_raw_dat[2:4], 
               filename = "full_corr_table_breaks_v_prod.doc", 
               table.number = 1, 
               show.sig.stars = TRUE, 
               landscape = TRUE)

apa.cor.table( breaks_dat, 
               filename = "short_corr_table_breaks_v_prod.doc", 
               table.number = 1, 
               show.sig.stars = TRUE, 
               landscape = TRUE)

```

### Assumptions, Figures, and Plots - Raw Data
Descriptives section already checked normality so here we need to focus on linearity and homoscedasticity. We're  
going to center the data outright because I don't want to run everything multiple times.

```{r assumptions_and_figures_raw}
# Centering the data
breaks_dat$length_centered <- breaks_dat$length - mean(breaks_dat$length)


# Check Scatterplot of Centered data: Break Length is PV, Productivity is OV
breaks_raw_scatter <- ggplot(breaks_dat, aes(length, product) )

breaks_raw_scatter + 
    geom_point() + 
    geom_smooth(method = "lm", colour = "Red") + 
    ggtitle("Break Length Relation to Employee Productivity") + 
    labs(x = "Break Length (m)", y = "Employee Productivity (0-100)") +
    theme_minimal() +
    theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
           axis.title.x = element_text(color = font_color),
           axis.title.y = element_text(color = font_color),
           axis.text.x = element_text(color = font_color),
           axis.text.y = element_text(color = font_color),
           legend.title = element_text(color = font_color),
           legend.text = element_text(color = font_color),
           panel.grid.minor = element_line(color = grid_color_minor),
           panel.grid.major = element_line(color = grid_color_major),
           panel.background = element_rect(fill = back_color, color = font_color)
         )

breaks_cent_scatter <- ggplot(breaks_dat, aes(length_centered, product) )

breaks_cent_scatter + 
    geom_point() + 
    geom_smooth(method = "lm", colour = "Red") + 
    ggtitle("Break Length Relation to Employee Productivity") + 
    labs(x = "Break Length - Centered (m)", y = "Employee Productivity (0-100)") +
    theme_minimal() +
    theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
           axis.title.x = element_text(color = font_color),
           axis.title.y = element_text(color = font_color),
           axis.text.x = element_text(color = font_color),
           axis.text.y = element_text(color = font_color),
           legend.title = element_text(color = font_color),
           legend.text = element_text(color = font_color),
           panel.grid.minor = element_line(color = grid_color_minor),
           panel.grid.major = element_line(color = grid_color_major),
           panel.background = element_rect(fill = back_color, color = font_color)
         )


# Check Residuals plot (Heteroscedasticity)
# To compute residuals we need to define the model as productivity regressed onto break length in a linear fashion
breaks_model_lin <- lm(product ~ length, data = breaks_dat)

# Fitted values vs. residuals to examine homoscedasticity
# NOTE use of .resid to plot residuals
breaks_resid_fig = ggplot( breaks_model_lin, aes(.fitted, .resid) )

breaks_resid_fig + 
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Productivity Values (0 - 100)") +
    ylab("Productivity Residuals (0 - 100)") +
    ggtitle("Residual vs. Fitted Plot") +
    theme_minimal() +
    theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
           axis.title.x = element_text(color = font_color),
           axis.title.y = element_text(color = font_color),
           axis.text.x = element_text(color = font_color),
           axis.text.y = element_text(color = font_color),
           legend.title = element_text(color = font_color),
           legend.text = element_text(color = font_color),
           panel.grid.minor = element_line(color = grid_color_minor),
           panel.grid.major = element_line(color = grid_color_major),
           panel.background = element_rect(fill = back_color, color = font_color)
         )

# I also want to see a residuals plot in multiples of SD of productivity to help spot outliers
breaks_resid_sd_fig = ggplot( breaks_model_lin, aes(.fitted , .resid / breaks_descr$descriptives$asDF$`product[sd]` ) )

breaks_resid_sd_fig + 
    geom_point(col = font_color) +
    geom_hline(yintercept=0, col="green3", linetype="dashed") +
    xlab("Fitted Productivity Values") +
    ylab(expression( "Residuals - Standardized (Multiples of " * sigma * ")") )  +
    ggtitle("Productivity Residual vs. Fitted Plot") +
    theme_minimal() +
    theme( plot.title = element_text(size = 12, face = "italic", color = font_color),
           axis.title.x = element_text(color = font_color),
           axis.title.y = element_text(color = font_color),
           axis.text.x = element_text(color = font_color),
           axis.text.y = element_text(color = font_color),
           legend.title = element_text(color = font_color),
           legend.text = element_text(color = font_color),
           panel.grid.minor = element_line(color = grid_color_minor),
           panel.grid.major = element_line(color = grid_color_major),
           panel.background = element_rect(fill = back_color, color = font_color)
         )  

# Run a Breusch Pagan Test of homoscedasticity
car::ncvTest(breaks_model_lin)
```

### Data Cleaning - OPTIONAL
If there are data issues that need to be adjusted, clean the data here.

``` {r data_cleaning_raw}

```

### Build Simple Regression Model
Centered data is already saved from scatter plots above (shift the y-intercept to the mean of the dataset).  
Now we can build the regression model.  

```{r simple_regression}
#simple regression looking at self-efficacy (centered) as a predictor of desire to lead in a group (e.g. self-efficacy (centered) regressed on desire to lead)
breaks_model_lin_reg <- linReg( data = breaks_dat, 
                                dep = 'product', 
                                covs = c('length_centered'), 
                                blocks = list( c( 'length_centered') ), 
                                modelTest = TRUE, 
                                stdEst = TRUE,
                                ci = TRUE)
breaks_model_lin_reg


```

### APA Tables

Practice outputting APA tables to a word doc just to see if that's easier than your notepad++ workflow
```{r apa_tables}
# We'll define a new linear model using the centered data.
breaks_model_lin_centered <- lm(product ~ length_centered, data = breaks_dat)

# Use reg.table method in APA package to generate regression table
apa.reg.table( breaks_model_lin_centered, filename="prod_regressed_to_breaks.doc", table.number = 2 )

```

