---
title: "PSYCH308A - Data Analysis 5 (DA5)"
author: "Brady C. Jackson"
date: "2024/10/15"

# Write document output to HTML, Word, and PDF output types with a Table of
# contents included.
output: 
  html_document:
    toc: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex

# This option here enables output to both HTML and PDF formats
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

```{r package_loading, message=FALSE, warning=FALSE}

# Load packages. Set messages and warnings to FALSE so I don't have to see the
# masking messages in the output.
library(psych)
library(jmv)       # for descriptive
library(ggplot2)
library(dplyr)
library(magrittr)
library(stringr)   # for sub_str
library(AER)

```

# Data Prep

First we need to load the data

```{r data_prep}

# Load the data as a dataframe
reading_dat <- read.csv("./DA5a.csv")

# Assert column names to be lowercase
colnames(reading_dat) <- tolower(colnames(reading_dat))

# Check the structure of the data:
print(head(reading_dat))
cat("\n\n")
str(reading_dat)

# Check uniqueness of fields of interrest for upcoming questions 
# (sex, education, and employment). We want to make sure that there are not 
# equal value inputs that would look different due to things like case or type:
# (e.g. High school graduate and HIGH School Graduatre and High School Complete)
uni_edu = unique(reading_dat$education) 
uni_sex = unique(reading_dat$sex)
uni_emp = unique(reading_dat$employment)

# Print the values for inspection
cat("\nUnique values in Education:\n") 
cat(uni_edu, sep="\n")
cat("---\n\n")

cat("\nUnique values in Sex:\n") 
cat(uni_sex, sep="\n")
cat("---\n\n")

cat("\nUnique values in Employment:\n") 
cat(uni_emp, sep="\n")
cat("---\n\n")
```

# Question 06

Data for Employment and Sex look appropriately unique. So we'll create a  
visualization (bar chart) for employment sorted by sex. Then we will print  
contingency tables and run the test  
  
## Q06 Visualization

Code below will create two bar-charts, one with the breakdown named by the   
original employment categories in the data, and one with the same data but  
labeled by more readable display names. Both are printed so that they can  
be inspected to ensure names were mapped correctly.  
  
```{r question_06_vis}

# Create a mapping vector setting each long-name to a shortened display name
emp_disp_name_map = c( "Employed full-time" = "Full-time",
                       "Employed part-time" = "Part-time",
                       "Have own business/self-employed" = "Self-employed",
                       "Not employed for pay" = "Not employed",
                       "Retired" = "Retired",
                       "Student" = "Student",
                       "Disabled" = "Disabled"
                      )

# Bar chart visualization. This code constructs a bar chart that shows each 
# employment type broken down by percentage of females vs. percentage of
# males at that level. Note: Two versions of the same chart are produced, one 
# uses default education level names and one with shortened display names. This  
# is just to ensure I mapped the display names correctly.
ggplot(reading_dat, aes( x=employment, fill=as.factor(sex) ) ) + 
    geom_bar(position = "fill") + 
    scale_fill_brewer(palette = "Set2") + 
    labs(y = "Percent", x = "Employment Type", fill= "Sex", title = "Employment by Sex") +
    theme_minimal() + 
    theme( axis.text.x = element_text(angle = 45, hjust = 1) )

ggplot(reading_dat, aes( x=employment, fill=as.factor(sex) ) ) + 
    geom_bar(position = "fill") + 
    scale_fill_brewer(palette = "Set2") + 
    scale_x_discrete(labels = emp_disp_name_map) +
    labs(y = "Percent", x = "Employment Type", fill= "Sex", title = "Employment by Sex") +
    theme_minimal() + 
    theme( axis.text.x = element_text(angle = 45, hjust = 1) )

```

## Q06 Contingency Table

```{r question_06_table}
# We create a table for consumption
emp_v_sex_tab <- prop.table( xtabs(~ sex +  employment, data = reading_dat), 1 )
round(emp_v_sex_tab, 2)

```

## Q06 NHST

```{r question_06_nhst}
jmv::contTables(data = reading_dat, 
                rows="sex", cols="employment", 
                exp=TRUE, 
                phiCra=TRUE
                )
```

# Question 07

Data for Education and Sex look appropriately unique. So we'll create a  
visualization (bar chart) for education sorted by sex. Then we will print  
contingency tables and run the test  

## Q07 Visualization

```{r question_07_vis}

# Create a mapping vector setting each long-name to a shortened display name
edu_disp_name_map = c( "Post-graduate training/professional school after college" = "Post Graduate",
                       "Some college, no 4-year degree" = "Some College",
                       "College graduate" = "College Graduate",
                       "High school graduate" = "HS Graduate",
                       "High school incomplete" = "HS Incomplete",
                       "Technical, trade or vocational school AFTER high school" = "Vocational after HS",
                       "None" = "None"
                      )

# Bar chart visualization. This code constructs a bar chart that shows each 
# level of education broken down by percentage of females vs. percentage of
# males at that level. Note: Two versions of the same chart are produced, one 
# default education level names and one with shortened display names. This is 
# just to ensure I mapped the display names correctly.
ggplot(reading_dat, aes( x=education, fill=as.factor(sex) ) ) + 
    geom_bar(position = "fill") + 
    scale_fill_brewer(palette = "Set2") + 
    labs(y = "Percent", x = "Education Level", fill= "Sex", title = "Education by Sex") +
    theme_minimal() + 
    theme( axis.text.x = element_text(angle = 45, hjust = 1) )

ggplot(reading_dat, aes( x=education, fill=as.factor(sex) ) ) + 
    geom_bar(position = "fill") + 
    scale_fill_brewer(palette = "Set2") + 
    scale_x_discrete(labels = edu_disp_name_map) +
    labs(y = "Percent", x = "Education Level", fill= "Sex", title = "Education by Sex") +
    theme_minimal() + 
    theme( axis.text.x = element_text(angle = 45, hjust = 1) )

```

## Q07 Contingency Table

```{r question_07_table}
# We create a table for consumption
edu_v_sex_tab <- prop.table( xtabs(~ sex + education, data = reading_dat), 1 )
round(edu_v_sex_tab, 2)

```

## Q07 NHST

```{r question_07_nhst}
jmv::contTables(data = reading_dat, 
                rows="sex", cols="education", 
                exp=TRUE, 
                phiCra=TRUE
                )

```

# Question 08

R was not used in any capacity for this question.
