---
title: "Correlation and Simple Regression Demo"
author: "Jessica Diaz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(car)
library(psych)
library(jmv)
library(ggplot2)
library(tidyverse)
library(magrittr)
library(praise)
library(apaTables)
library(broom)
```

**Example 1**
Using participants who self-identified as active in one or more Reddit groups, this study examined the relationships between several group and individual variables, individuals desire to lead, and their actual decision to pursue a leadership opportunity. 

Variables:

-Participant: ID Number
-Group ID: Level of Group Identification (Low/High)
-Support: Level of Support from Group (Low/High)
-Group.Support: Level of Support from the Group (Continuous)
-Desire to Lead: Desire to lead the group
-Self Efficacy: Belief in their ability to lead the group
-MTL Affective: General affective motivation to lead ("I like to be in charge")
-MTL Socio Normative: General socio-normative to lead ("I will lead when asked")
-Dominance: Influence via threat/power
-Prestige: Influence via specialist knowledge
-Leader Identity: Degree to which they view themselves as a leader
-LdrBehavResponse: Decision to Lead (Yes/No)
-Age: Age
-Gender: Gender

*Run the appropriate analyses and report relevant statistics in APA format. Please interpret the results for both someone who is knowledgable in statistics and for someone who is not.*

```{r}
#Get some praise to start the day
praise()
```

```{r}
#Read in your data. As long as its in the same folder as the RMD file, this is the only code you need. 
dat <- read.csv("lead.csv")

#make the column names lowercase
colnames(dat) %<>% tolower()

#Take a quick glimpse at your data to see how its categorized (or you can open the data from your environment)
glimpse(dat)
```

```{r}
#descriptives of your data
desc <- descriptives(data = dat[-c(1:3,12:13,15)], hist = TRUE, sd = TRUE, range = TRUE, skew = TRUE, kurt = TRUE)
desc

#if you have a lot of variables, it can be helpful to run descriptives using psych's describe function rather than jmv's descriptives function because it organizes them vertically.
desc2 <- describe(dat[-c(1:3,12:13,15)])
desc2
```

```{r}
#all of the Correlation Options

#Option 1 (cor function in stats package --> ask it to round to 2 decimal plances)
cor <- cor(dat[c(4:8)]) %>% round(2)
cor

#make it visual
library(corrplot)
corrplot(cor, method="circle")
corrplot(cor, method="color", type = "lower")
corrplot(cor, method="number", type = "lower")

#Option 2 (corrMatrix function in jmv package) - this is my preferred function!
cor2 <- corrMatrix(dat[-c(1:3,12:13,15)], flag = TRUE)
cor2

#Option 3 (APA Table using the apaTables package)
#note that the confidence interval argument has been depricated and they must be manually removed if not desired
apa.cor.table(dat[-c(1:3,12:13,15)], filename = "MTL Correlation Table.doc", table.number = 1, show.sig.stars = TRUE, landscape = TRUE)
```

**Self-Efficacy**
```{r}
#create a scatter plot of self efficacy and desire to lead to confirm linearity
scatter.se <- ggplot(dat, aes(self.efficacy, desire.to.lead))

#make that scatterplot relationship visually appealing
scatter.se + geom_point() + geom_smooth(method = "lm", colour = "Red") + ggtitle("Self-Efficacy Predicting Desire to Lead") + labs(x = "Self Efficacy", y = "Desire to Lead") + theme(plot.title = element_text(hjust = 0.5))
```

```{r}
model.se <- lm(desire.to.lead~self.efficacy, data = dat)

#fitted values vs. residuals to examine homoscedasticity
# NOTE use of .resid to plot residuals
ggplot(model.se, aes(.fitted, .resid))+geom_point()+geom_hline(yintercept=0, col="green", linetype="dashed")+xlab("Fitted Values")+ylab("Residuals")+ggtitle("Residual vs. Fitted Plot")+theme_bw()

#Breusch Pagan Test of homoscedasticity
ncvTest(model.se)
```

***Simple Regression***
```{r}
#simple regression looking at self-efficacy as a predictor of desire to lead in a group (e.g. self-efficacy regressed on desire to lead)
# stdEst gives us Beta
model.se.uc <- linReg(data = dat, 
                 dep = 'desire.to.lead', 
                 covs = c('self.efficacy'), #all the predictors you want available
                 blocks = list(c('self.efficacy')), 
                 modelTest = TRUE, 
                 stdEst = TRUE,
                 ci = TRUE)
model.se.uc
```

***Same Simple Regression but Centered***
```{r}
#center your predictors for interpretability
dat$self.efficacy.c <- dat$self.efficacy - mean(dat$self.efficacy)

#simple regression looking at self-efficacy (centered) as a predictor of desire to lead in a group (e.g. self-efficacy (centered) regressed on desire to lead)
model.se.c <- linReg(data = dat, 
                 dep = 'desire.to.lead', 
                 covs = c('self.efficacy.c'), 
                 blocks = list(c('self.efficacy.c')), 
                 modelTest = TRUE, 
                 stdEst = TRUE,
                 ci = TRUE)
model.se.c
```

```{r}
#APA Regression Table - not perfect, but a start if you want it

#specify model using regular lm function first
model.se <- lm(desire.to.lead ~ self.efficacy.c, data = dat)
#generate APA regression table
apa.reg.table(model.se,filename="se_on_dtl.doc", table.number=2)
```

**Leader Identity**
```{r}
#create a scatter plot of leader identity and desire to lead
scatter.li <- ggplot(dat, aes(leader.identity, desire.to.lead))

#make that scatterplot relationship visually appealing
scatter.li + geom_point() + geom_smooth(method = "lm", colour = "Red") + ggtitle("Leader Identity Predicting Desire to Lead") + labs(x = "Leader Identity", y = "Desire to Lead") + theme(plot.title = element_text(hjust = 0.5))
```

```{r}
model.li <- lm(desire.to.lead~leader.identity, data = dat)

#fitted values vs. residuals to examine homoscedasticity
ggplot(model.li, aes(.fitted, .resid))+geom_point()+geom_hline(yintercept=0, col="green", linetype="dashed")+xlab("Fitted Values")+ylab("Residuals")+ggtitle("Residual vs. Fitted Plot")+theme_bw()

#Breusch Pagan Test of homoscedasticity
ncvTest(model.li)
```

```{r}
#center your predictors for interpretability
dat$leader.identity.c <- dat$leader.identity - mean(dat$leader.identity)

#simple regression looking at leader identity (centered) as a predictor of desire to lead in a group (e.g. leader identity (centered) regressed on desire to lead)
model.li.c <- linReg(data = dat, 
                 dep = 'desire.to.lead', 
                 covs = c('leader.identity'), 
                 blocks = list(c('leader.identity')), 
                 modelTest = TRUE, 
                 stdEst = TRUE,
                 ci = TRUE)
model.li.c
```

**Group Support**
```{r}
#create a scatter plot of group support and desire to lead
scatter.gs <- ggplot(dat, aes(group.support, desire.to.lead))

#make that scatterplot relationship visually appealing
scatter.gs + geom_point() + geom_smooth(method = "lm", colour = "Red") + ggtitle("Group Support Predicting Desire to Lead") + labs(x = "Group Support", y = "Desire to Lead") + theme(plot.title = element_text(hjust = 0.5))
```

```{r}
model.gs <- lm(desire.to.lead~group.support, data = dat)

#fitted values vs. residuals to examine homoscedasticity
ggplot(model.gs, aes(.fitted, .resid))+geom_point()+geom_hline(yintercept=0, col="green", linetype="dashed")+xlab("Fitted Values")+ylab("Residuals")+ggtitle("Residual vs. Fitted Plot")+theme_bw()

#Breusch Pagan Test of homoscedasticity
ncvTest(model.gs)
```

```{r}
#center your predictors for interpretability
dat$group.support.c <- dat$group.support - mean(dat$group.support) 

#simple regression looking at group support (centered) as a predictor of desire to lead in a group (e.g. group support  (centered) regressed on desire to lead)
model.gs.c <- linReg(data = dat, 
                 dep = 'desire.to.lead', 
                 covs = c('group.support'), 
                 blocks = list(c('group.support')), 
                 modelTest = TRUE, 
                 stdEst = TRUE,
                 ci = TRUE)
model.gs.c
```


